---
title: "PRISM"
author: "AnnaClaire Marley"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analysis of PRISM data

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(tidyquant)
library(dataRetrieval)
library(measurements)
```

```{r, echo = FALSE}
# path to source graphing functions
path_graph <- "../Functions/Graphing/"
```

```{r, echo = FALSE}
# source functions
source(paste0(path_graph, "graph_season_trend.R"))
```

```{r, echo = FALSE}
# for organizing graphs in month order of water years
wy_order = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")
```



### Precipitation

- precipitation needs to be made into volume falling on the Chuskas
- Since this data is already a sum of all the grid cells of precipitaiton falling on the Chuskas we need to multiply this sum by area of each PRISM pixel (16 km2)

Daily
```{r, echo = FALSE}
prism_pixel_m2 = conv_unit(16, "km2", "m2")

# read in prism data -- sum of precipiation across the chuska mountains
chuska_precip_day <- read_csv("../Data/chuska_prism_precip.csv", col_names = c("Date", "precip_mm"), skip = 1) %>% 
  mutate(Date = ymd(Date)) %>% 
  mutate(precip_m3 = conv_unit(precip_mm, "mm", "m")*prism_pixel_m2)

# initial look
ggplot(chuska_precip_day, aes(x = Date, y = precip_mm)) +
  geom_line()

```


Monthly sum
```{r, echo = FALSE}
chuska_precip_month <- chuska_precip_day %>% 
  mutate(year = year(Date)) %>% 
  mutate(month = month(Date)) %>% 
  group_by(year, month) %>% 
  summarize(precip_mm = sum(precip_mm)) %>% 
  mutate(Date = ymd(paste0(year, "-", month, "-01"))) %>% 
  addWaterYear() %>% 
  ungroup()

ggplot(chuska_precip_month, aes(x = Date, y = precip_mm)) +
  geom_line() +
  geom_ma(ma_fun = SMA, n = 12, color = "red", linetype = 5) +
  labs(
    title = "Monthly sum of precipitation Chuska Mountains"
  ) +
  theme_classic()

ggplot(chuska_precip_month, aes(x = waterYear, y = precip_mm)) +
  geom_col() +
  facet_wrap(~month)
```


```{r, echo = FALSE}

# set the seasons
winter = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr")
summer = c("May", "Jun", "Jul", "Aug", "Sep")
winter_wet = c("Dec", "Jan", "Feb", "Mar")
spring_dry =  c("Apr", "May", "Jun")
summer_wet = c("Jul", "Aug", "Sep")


# Yearly chuska total precipitation m3
chuska_precip_wy <- chuska_precip_day %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  summarize(wy_precip_m3 = sum(precip_m3))


# show how much total precipitation this area gets every month over the years
prism_precip_months <- chuska_precip_day %>% 
  mutate(year = year(Date), month = month(Date, label=TRUE,abbr=TRUE)) %>% 
  addWaterYear() %>% 
  group_by(waterYear, month) %>% 
  summarize(
    monthly_precip_mm = sum(precip_mm)
  ) %>% 
  mutate(bimodal_season = ifelse(month %in% summer, "summer",
                                 "winter")) %>% 
  mutate(wet_dry = ifelse(month %in% winter_wet, "winter_wet",
                          ifelse(month %in% spring_dry, "spring_dry",
                                 ifelse(month %in% summer_wet, "summer_wet",
                                        "fall_dry"))))
  

# plot monthly total precip
ggplot(prism_precip_months, aes(x = waterYear, y = monthly_precip_mm)) +
  geom_col() +
  facet_wrap(~factor(month, wy_order)) +
  labs(
    x = "Water Year",
    y = "Total Precipitation (mm)",
    title = "Chuska Total PRISM Precipitation By Month"
  ) +
  theme_bw()

# show how much total precipitation this area gets every month over the years in m3
prism_precip_months_m3 <- chuska_precip_day %>% 
  mutate(year = year(Date), month = month(Date, label=TRUE,abbr=TRUE)) %>% 
  addWaterYear() %>% 
  group_by(waterYear, month) %>% 
  summarize(
    monthly_precip_m3 = sum(precip_m3)
  ) %>% 
  left_join(chuska_precip_wy, by = "waterYear") %>% 
  mutate(month_wy_fract = monthly_precip_m3/wy_precip_m3)

# plot monthly total precip m3
ggplot(prism_precip_months_m3, aes(x = waterYear, y = monthly_precip_m3)) +
  geom_col() +
  facet_wrap(~factor(month, wy_order)) +
  labs(
    x = "Water Year",
    y = "Total Precipitation (m3)",
    title = "Chuska Total PRISM Precipitation By Month"
  ) + 
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

# plot monthly total precip m3 as a fraction of yearly total
ggplot(prism_precip_months_m3, aes(x = waterYear, y = month_wy_fract)) +
  geom_col() +
  facet_wrap(~factor(month, wy_order)) +
  labs(
    x = "Water Year",
    y = "Fraction",
    title = "Chuska Monthly Fraction of Total PRISM Precipitation"
  ) + 
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()



```

```{r, echo = FALSE}
# plot total precip separated by winter/summer
prism_precip_months %>% 
  group_by(waterYear, bimodal_season) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
ggplot(aes(x = waterYear, y = season_precip_mm)) +
  geom_col(aes(fill = bimodal_season)) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Chuska Total PRISM Precipitation By Season",
    fill = "Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

# plot total precip separated by winter/summer fraction
prism_precip_months %>% 
  group_by(waterYear, bimodal_season) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
  mutate(percent = season_precip_mm/sum(season_precip_mm)) %>% 
ggplot(aes(x = waterYear, y = percent)) +
  geom_col(aes(fill = bimodal_season)) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Chuska Total PRISM Precipitation By Season",
    fill = "Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

# plot total precip separated by wet/dry seasons
prism_precip_months %>% 
  group_by(waterYear, wet_dry) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
ggplot(aes(x = waterYear, y = season_precip_mm)) +
  geom_col(aes(fill = wet_dry)) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Chuska Total PRISM Precipitation By Season",
    fill = "Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

# plot total precip separated by wet/dry seasons percent
prism_precip_months %>% 
  group_by(waterYear, wet_dry) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
  mutate(percent = season_precip_mm/sum(season_precip_mm)) %>% 
ggplot(aes(x = waterYear, y = percent)) +
  geom_col(aes(fill = wet_dry)) +
  labs(
    x = "Year",
    y = "Fraction of Total Precipitation",
    title = "Chuska Total PRISM Precipitation By Wet and Dry Season",
    fill = "Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()
```


### Temperature

```{r, echo = FALSE}
# read in mean temp data (C)
chuska_tmean <- read_csv("../Data/chuska_prism_tmean.csv") %>% 
  mutate(Date = ymd(Date)) %>% 
  arrange(Date) %>% 
  addWaterYear()

# initial look
ggplot(chuska_tmean, aes(x = Date, y = tmean)) +
  geom_line()

```

Number of freezing days, when average temperature is at or below 0

```{r, echo = FALSE}
# number of freezing days per water year
# remove 1981 because wasn't a full winter (data starts in Jan)
num_freeze <- chuska_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  group_by(waterYear) %>% 
  summarize(ndayfr=sum(freeze)) %>% 
  filter(waterYear != 1981)

# graphit
graph_season_trend(num_freeze, num_freeze$ndayfr, num_freeze$waterYear, 
                   title = "Number of Freezing Days per Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Number of Freezing Days")  

graph_season_trend(num_freeze, num_freeze$ndayfr, num_freeze$waterYear, 
                   title = "Number of Freezing Days per Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Number of Freezing Days",
                   graph_type = "point") 
```

Temperature of hottest average temperature days each water year

```{r, echo = FALSE}
# hottest day each water year
chuska_hot <- chuska_tmean %>% 
  group_by(waterYear) %>% 
  summarize(tmean = max(tmean)) %>% 
  filter(waterYear != 2020) # not a complete water year

# look at dates of hottest days
chuska_hot_winter_days <- chuska_tmean %>% 
  filter(month(Date) %in% c(11,12,1,2,3,4)) %>% 
  mutate(month_name = month(Date, label = TRUE, abbr = TRUE)) %>% 
  group_by(waterYear) %>% 
  top_n(1, tmean) %>% 
  mutate(water_start_date = mdy(paste0("oct", "-", "1", "-", year(Date)-1))) %>% 
  mutate(day_wy = Date-water_start_date) %>% 
  mutate(day_wy = ifelse(month_name == "Nov", 35, day_wy))

# histogram of day of water year
ggplot(chuska_hot_winter_days, aes(x = day_wy)) +
  geom_histogram() +
  labs(
    title = "Histogram of Day of Water of Year Hottest Day of Winter Occurs",
    x = "Day of Water Year"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

graph_season_trend(chuska_hot_winter_days, chuska_hot_winter_days$day_wy,
                   chuska_hot_winter_days$waterYear, 
                   title = "Day of Water Hottest Day of Winterr", 
                   xlabel = "Water Year",
                   ylabel = "Day of Water Year") 

# graphit
graph_season_trend(chuska_hot, chuska_hot$tmean, chuska_hot$waterYear, 
                   title = "Temperature of Hottest Day Each Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Temperature C")  

# hottest day each winter water year winter = nov, dec, jan, feb, mar, apr
chuska_hot_winter <- chuska_tmean %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11, 12, 1, 2, 3, 4)) %>% 
  group_by(waterYear) %>% 
  summarize(tmean = max(tmean))

# graphit
graph_season_trend(chuska_hot_winter, chuska_hot_winter$tmean, chuska_hot_winter$waterYear, 
                   title = "Average Temperature of Hottest Day Each Winter", 
                   xlabel = "Winter Water Year",
                   ylabel = "Temperature C")  

ggplot(chuska_hot_winter, aes(x = waterYear, y = tmean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# average temperature winter water year winter = nov, dec, jan, feb, mar, apr
chuska_av_winter <- chuska_tmean %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11, 12, 1, 2, 3, 4)) %>% 
  group_by(waterYear) %>% 
  summarize(tmean = mean(tmean))

# graphit
graph_season_trend(chuska_av_winter, chuska_av_winter$tmean, chuska_av_winter$waterYear, 
                   title = "Average Temperature Each Winter", 
                   xlabel = "Winter Water Year",
                   ylabel = "Temperature (°C)",
                   graph_type = "point")  

```



## Temp and precip

```{r, echo = FALSE}
# combine temp and snow 
ch_tav_precip <- left_join(chuska_tmean, chuska_precip_day, by = "Date") %>% 
  mutate(precip_type = ifelse(tmean <= 0, "snow", "rain"))
```


**Amount of snow each water year**


```{r, echo = FALSE}
# snotel data is accumulated precipitation, so we need to sum prism snowfall to estimate how much snow
chuska_snow <- ch_tav_precip %>% 
  mutate(month = month(Date)) %>% 
  filter(precip_type != is.na(precip_type)) %>% 
  mutate(prism_snow_mm = ifelse(precip_type == "snow", precip_mm, 0)) %>% 
  mutate(prism_snow_m3 = ifelse(precip_type == "snow", precip_m3, 0)) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(4,5,6,7,8,9,10), 0, 
                                      prism_snow_mm_accum)) %>% 
  mutate(prism_snow_m3_accum = cumsum(prism_snow_m3)) %>%  # accumulate prism precip
  mutate(prism_snow_m3_accum = ifelse(month(Date) %in% c(4,5,6,7,8,9,10), 0, 
                                      prism_snow_m3_accum))
```

Is more preciptiation falling as rain or snow in winter?

```{r, echo = FALSE}
chuska_month_precip_all <- chuska_snow %>% 
  ungroup() %>% 
  mutate(month_name = month(Date, abbr = TRUE, label = TRUE)) %>% 
  group_by(waterYear, month_name, precip_type) %>% 
  summarize(precip_mm = sum(precip_mm, na.rm = TRUE)) %>% 
  mutate(fraction_precip = precip_mm/sum(precip_mm))

chuska_month_precip_all %>% 
  filter(month_name %in% c("Nov", "Dec", "Jan", "Feb", "Mar", "Apr")) %>% 
ggplot(aes(x = waterYear, y = precip_mm)) +
  geom_col(aes(fill = precip_type)) +
  facet_wrap(~factor(month_name, wy_order)) +
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    fill = "Precipitation Type"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()

chuska_month_precip_all %>% 
  filter(month_name %in% c("Nov", "Dec", "Jan", "Feb", "Mar", "Apr")) %>% 
  filter(precip_type == "snow") %>% 
ggplot(aes(x = waterYear, y = fraction_precip)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~factor(month_name, wy_order)) +
  labs(
    x = "Water Year",
    y = "Fraction",
    title = "Fraction of Total Monthly Precipitation Falling as Snow in the Chuskas"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()

jan_snow_fraction <- chuska_month_precip_all %>% 
  filter(month_name == "Jan") %>% 
  filter(precip_type == "snow")

summary(lm(jan_snow_fraction$fraction_precip ~ jan_snow_fraction$waterYear))


chuska_wint_precip_fract <- chuska_snow %>% 
  ungroup() %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  group_by(waterYear, precip_type) %>% 
  summarize(precip_mm = sum(precip_mm, na.rm = TRUE)) %>% 
  mutate(fraction_precip = precip_mm/sum(precip_mm))


snow_fraction <- chuska_wint_precip_fract %>% 
  filter(precip_type == "snow")

graph_season_trend(snow_fraction, snow_fraction$fraction_precip, snow_fraction$waterYear,
                    xlabel = "Water Year",
                   ylabel = "Fraction",
                   title = "Fraction of Total Winter Precipitation Falling as Snow",
                   graph_type = "point")

```

Is less snow falling in the Spring?

```{r, echo = FALSE}
ch_snow_spring <- chuska_month_precip_all %>% 
  filter(month_name %in% c("Mar", "Apr")) %>% 
  filter(precip_type == "snow")

ggplot(ch_snow_spring, aes(x = waterYear, y = precip_mm)) +
  geom_col() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~month_name) +
  labs(
    x = "Water Year",
    y = "Snowfall (mm)",
    title = "PRISM Estimated Total Snowfall Each Month in the Chuskas"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()

ch_snow_march <- ch_snow_spring %>% 
  filter(month_name == "Mar")

graph_season_trend(ch_snow_march, ch_snow_march$precip_mm, ch_snow_march$waterYear,
                   title = "March Total Precipitation Chuskas",
                   xlabel = "Water Year",
                   ylabel = "Estimated Total Snowfall (mm)")

summary(lm(ch_snow_march$precip_mm ~ ch_snow_march$waterYear))

```



```{r, echo = FALSE}
# prism snow accumulation
ggplot(chuska_snow, aes(x = Date, y = prism_snow_mm_accum/1000)) +
  geom_line() +
  labs(
    x = "Year",
    y = "PRISM Snow Accumulation (m)"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

# total amount of snow that fell on the chuskas
ch_snow_max <- chuska_snow %>% 
  filter(waterYear > 1981) %>% 
  summarize(snow_mm = max(prism_snow_mm_accum, na.rm = TRUE)) %>% 
  mutate(prism_snow_m_accum = conv_unit(snow_mm, "mm", "m"))

# plot it
graph_season_trend(ch_snow_max, (ch_snow_max$snow_mm)/1000, ch_snow_max$waterYear, 
                   title = "Max Accumulated Snowfall", 
                   xlabel = "Water Year",
                   ylabel = "Snow (m)") 

# total amount of snow that fell on the chuskas in m3
ch_snow_max_m3 <- chuska_snow %>% 
  filter(waterYear > 1981) %>% 
  summarize(snow_m3 = max(prism_snow_m3_accum, na.rm = TRUE)) 

ggplot(ch_snow_max_m3, aes(x = waterYear, y = snow_m3)) +
  geom_col() +
  labs(
    x = "Water Year",
    y = bquote("Maximum Accumulated Snowfall " (m^3))
  ) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()



# anomaly for max accumulated snowfall PRISM
ch_snow_max_mean_82_11 <- ch_snow_max %>% 
  filter(waterYear <= 2011) %>% 
  summarize(mean = mean(prism_snow_m_accum))
ch_snow_max_sd_82_11 <- ch_snow_max %>% 
  filter(waterYear <= 2011) %>% 
  summarize(sd = sd(prism_snow_m_accum))

ch_snow_max_anom <- ch_snow_max %>% 
  mutate(mean_82_11 = ch_snow_max_mean_82_11$mean) %>% 
  mutate(sd_82_11 = ch_snow_max_sd_82_11$sd) %>% 
  mutate(anomaly = (prism_snow_m_accum-mean_82_11)/sd_82_11) %>% 
  mutate(sign = ifelse(anomaly < 0, "negative", "positive"))

ggplot(ch_snow_max_anom, aes(x = waterYear, y = anomaly)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("red", "darkgreen")) +
  labs(
    x = "Water Year",
    y = "Anomaly",
    title = "Max PRISM Accumulated Snowfall Anomaly From 1982-2011 Average"
  ) +
  theme_classic()

```

#### Observation

- appears to be no trend in max accumulated snowfall each winter using PRISM

**Nunber of snow days**

```{r, echo = FALSE}
# snow days, # get rid of 1981 because not full water year
snow_days <- ch_tav_precip %>% 
  filter(precip_mm > 0) %>% 
  group_by(waterYear) %>% 
  filter(waterYear != 1981) %>% 
  count(precip_type) %>% 
  mutate(perc = (n/sum(n))*100)

# plot number of days
snow_days %>% 
filter(waterYear != 2020) %>%  
ggplot(aes(x = waterYear, y = n)) +
  geom_col(aes(fill = precip_type)) +
  labs(
    x = "Number of Days",
    y = "Water Year",
    fill = "Precipitation Type",
    title = "Number of Days Each Water Year Falling as Rain or Snow"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

# plot percent of days
snow_days %>% 
filter(waterYear != 2020) %>% 
ggplot(aes(x = waterYear, y = perc)) +
  geom_col(aes(fill = precip_type)) +
  labs(
    x = "Water Year",
    y = "Percent of Days",
    fill = "Precipitation Type",
    title = "Percent of Days Each Water Year Falling as Rain or Snow"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

# winter snow days
snow_days_win <- ch_tav_precip %>% 
  filter(precip_mm > 0) %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  group_by(waterYear) %>% 
  filter(waterYear != 1981) %>% 
  count(precip_type) %>% 
  mutate(perc = (n/sum(n))*100)

snow_days_win %>% 
ggplot(aes(x = waterYear, y = perc)) +
  geom_col(aes(fill = precip_type)) +
  labs(
    x = "Water Year",
    y = "Percent of Days",
    fill = "Precipitation Type",
    title = "Percent of Precipitation Days Each Winter Falling as Rain or Snow"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()
```


















