---
title: "Snotel"
author: "AnnaClaire Marley"
date: "6/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This RMarkdown conducts initial time series analyses on two SNOTEL sites in the Chuska Mountains located in the Navajo Nation.

**Read in packages**

```{r, message=FALSE, include = FALSE}
library(RNRCS)
library(tidyverse)
library(lubridate)
library(gridExtra)
library(measurements)
library(mltools)
library(dataRetrieval)
library(naniar)
```

```{r, include = FALSE}
# set path to source funcitons from data cleaning folder
path_clean = "../Functions/Cleaning/"

```

```{r, echo = FALSE}
# source functions
source(paste0(path_clean, "to_water_week.R")) # for calculating week of water year
```



### Navajo Whiskey Creek

* Station ID:	1138

* State:	New Mexico

* Network:	SNOTEL

* County:	San Juan

* Elevation:	9050 ft.

* Latitude:	36.18

* Longitude:	-108.95

* HUC:	140802040206

```{r, echo = FALSE}
# fixes numbers showing up in scientific notation
options(scipen = 999)

# grab snotel data from NRCS
whiskey_creek_snotel <- grabNRCS.data(network = "SNTL",
                                      site_id = 1138,
                                      timescale = "daily",
                                      DayBgn = "1984-10-31",
                                      DayEnd = "2020-06-01")

# clean dataset
wc_clean <- whiskey_creek_snotel %>% 
  select(Date,
          "air_temp_av_F" = Air.Temperature.Average..degF., 
         "air_temp_max_F" = Air.Temperature.Maximum..degF., 
        "air_temp_min_F" =  Air.Temperature.Minimum..degF.,
        "precip_in" = Precipitation.Accumulation..in..Start.of.Day.Values,
        "snow_depth_in" = Snow.Depth..in..Start.of.Day.Values,
        "swe_in" = Snow.Water.Equivalent..in..Start.of.Day.Values) %>% 
  mutate(Date = ymd(Date)) %>% 
  mutate(snow_depth_in = ifelse(snow_depth_in < 100, 
                                snow_depth_in, NA)) # get rid of outliers

# convert everything to metric units
wc_clean_metric <- wc_clean %>% 
  mutate(air_temp_av_C = round(conv_unit(air_temp_av_F, "F", "C"),2)) %>% 
  mutate(air_temp_max_C = round(conv_unit(air_temp_max_F, "F", "C"),2)) %>% 
  mutate(air_temp_min_C = round(conv_unit(air_temp_min_F, "F", "C"),2)) %>% 
  mutate(precip_mm = conv_unit(precip_in, "inch", "mm")) %>% 
  mutate(snow_depth_mm = conv_unit(snow_depth_in, "inch", "mm")) %>%
  mutate(swe_mm = conv_unit(swe_in, "inch", "mm")) %>% 
  select(Date, air_temp_av_C, air_temp_max_C, air_temp_min_C, precip_mm, snow_depth_mm, swe_mm)
```

### First explore NAs
```{r, echo = FALSE}
# table summary of missing variables
miss_var_summary(wc_clean_metric)

# plot of missing variables
vis_miss(wc_clean_metric, sort = TRUE)

# visualize NA intersections across variables with an upset plot
gg_miss_upset(wc_clean_metric, 
              nsets = 6)
```

- The horizontal black bars in the lower left indicate the number of NAs total for each variable, with the variable name shown to the right of the corresponding horizontal bar

- The vertical black lines with black dots indicate the variables between which the frequency of co-occurring NAs is indicated by the vertical black bars


#### Daily basic trends

```{r, echo = FALSE}
# SWE
wc_swe_daily <- ggplot(wc_clean_metric, aes(x = Date, y = swe_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (mm)",
                      title = "Daily SWE"
                    ) +
                    theme_classic()

# snow depth
wc_depth_daily <- ggplot(wc_clean_metric, aes(x = Date, y = snow_depth_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (mm)",
                      title = "Daily Snow Depth"
                    ) +
                    theme_classic()

# max temp
wc_max_temp_daily <- ggplot(wc_clean_metric, aes(x = Date, y = air_temp_max_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Daily Maximum Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(wc_swe_daily, wc_depth_daily, wc_max_temp_daily, ncol = 2)

```



#### Monthly basic trends

```{r, echo = FALSE}
# maximum monthly data
wc_max_monthly <- wc_clean_metric %>% 
  group_by(year = year(Date), 
           month = month(Date)) %>% 
  summarize(air_temp_av_C = max(air_temp_av_C, na.rm = TRUE),
            air_temp_max_C = max(air_temp_max_C, na.rm = TRUE),
            air_temp_min_C = max(air_temp_min_C, na.rm = TRUE),
            swe_mm = max(swe_mm, na.rm = TRUE),
            snow_depth_mm = max(snow_depth_mm, na.rm = TRUE)) %>% 
  mutate(Date = ymd(paste0(year, "-", month, "-01")))

# SWE
wc_swe_monthly <- ggplot(wc_max_monthly, aes(x = Date, y = swe_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (mm)",
                      title = "Max Monthly SWE"
                    ) +
                    theme_classic()

# snow depth
wc_depth_monthly <- ggplot(wc_max_monthly, aes(x = Date, y = snow_depth_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (mm)",
                      title = "Max Monthly Snow Depth"
                    ) +
                    theme_classic()

# max temp
wc_max_temp_monthly <- ggplot(wc_max_monthly,
                              aes(x = Date, y = air_temp_max_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Maximum Air Temperature"
                    ) +
                    theme_classic()


# max average temp
wc_max_av_temp_monthly <- ggplot(wc_max_monthly,
                              aes(x = Date, y = air_temp_av_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Average Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(wc_swe_monthly, wc_depth_monthly, wc_max_temp_monthly, wc_max_av_temp_monthly, ncol = 2)


```

#### Water year trends
```{r, echo = FALSE}
# water year max values
wc_max_wy <- wc_max_monthly %>% 
  addWaterYear() %>% 
  ungroup() %>% 
  group_by(waterYear) %>% 
  summarize(
    swe_mm = max(swe_mm),
    snow_depth_mm = max(snow_depth_mm)
  ) %>% 
  filter(waterYear > 2009)

ggplot(wc_max_wy, aes(x = waterYear, y = swe_mm)) +
  geom_point() +
  geom_line()
```


### Compare to SNODAS data (extracted for the same lat/long as snotel site)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# read in snodas swe
wc_snodas_swe <- read_csv("../Data/whiskey_creek_snodas_swe.csv",
                          col_names = c("Date", "snodas_swe_mm"), 
                          skip = 1)
# read in snodas snow depth
wc_snodas_depth <- read_csv("../Data/whiskey_creek_snodas_depth.csv",
                          col_names = c("Date", "snodas_snow_depth_mm"), 
                          skip = 1)
```

#### Daily data

```{r, echo = FALSE}
# combine snodas and snotel
wc_snodas_snotel <- left_join(wc_snodas_swe, wc_snodas_depth) %>% 
  left_join(wc_clean_metric, by = "Date") %>% 
  select(Date, snodas_swe_mm,  snodas_snow_depth_mm,
         "snotel_depth_mm" = snow_depth_mm, 
         "snotel_swe_mm" = swe_mm) %>% 
  addWaterYear()

# tidy the data for plotting
wc_tidy_snodas_snotel <- wc_snodas_snotel %>% 
  pivot_longer(c(snodas_swe_mm, snodas_snow_depth_mm, 
                 snotel_depth_mm, snotel_swe_mm),
               names_to = "type", values_to = "value_mm") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
### SWE ###
wc_das_tel_swe <- wc_tidy_snodas_snotel %>% 
  filter(type == "snodas_swe_mm" | type ==  "snotel_swe_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "SWE (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL SWE Whiskey Creek Station"
  ) +
  theme_classic()
wc_das_tel_swe

# clean the data for graphing
wc_snodas_snotel_wint <- wc_snodas_snotel %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  filter(waterYear > 2009)# 2009 not a full water year
  
# all correlation    
  ggplot(wc_snodas_snotel_wint, aes(x = snotel_swe_mm, y = snodas_swe_mm)) +
    geom_point(aes(color = waterYear)) +
    geom_smooth(method = lm) +
    scale_x_continuous(expand = c(0,0)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(
      x = "SNOTEL SWE (mm)",
      y = "SNODAS SWE (mm)",
      title = "SNOTEL and SNODAS Daily SWE Relationship Nov-Apr"
    ) +
    theme_classic()

# correlations by water year
  ggplot(wc_snodas_snotel_wint, aes(x = snotel_swe_mm, y = snodas_swe_mm)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_wrap(~waterYear) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL SWE (mm)",
        y = "SNODAS SWE (mm)",
        title = "SNOTEL and SNODAS Daily SWE Relationship Nov-Apr"
      ) +
      theme_classic()
  
  # correlations for 2014 and 2018
wc_snodas_snotel_wint %>% 
  filter(waterYear %in% c(2018)) %>% 
   ggplot(aes(x = snotel_swe_mm, y = snodas_swe_mm)) +
      geom_point() +
      geom_smooth(method = lm, se = FALSE) +
      facet_wrap(~waterYear) +
   #   scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL SWE (mm)",
        y = "SNODAS SWE (mm)",
        title = "SNOTEL and SNODAS Daily SWE Relationship Nov-Apr"
      ) +
      theme_classic()
  
```

**stats**
```{r}
# correlation
cor.test(wc_snodas_snotel_wint$snodas_swe_mm, wc_snodas_snotel_wint$snotel_swe_mm)

# linear model 
summary(lm(wc_snodas_snotel_wint$snodas_swe_mm ~ wc_snodas_snotel_wint$snotel_swe_mm))

# root mean square error
rmse(wc_snodas_snotel_wint$snodas_swe_mm, wc_snodas_snotel_wint$snotel_swe_mm, na.rm = TRUE)
```

```{r, echo = FALSE}
# snow depth
wc_das_tel_depth <- wc_tidy_snodas_snotel %>% 
  filter(type == "snodas_snow_depth_mm" | type ==  "snotel_depth_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "Snow depth (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL Snow Depth"
  ) +
  theme_classic()
wc_das_tel_depth

# correlation
ggplot(wc_snodas_snotel_wint, aes(x = snotel_depth_mm, y = snodas_snow_depth_mm)) +
  geom_point(aes(color = waterYear)) +
  geom_smooth(method = lm) +
  labs(
    x = "SNOTEL Snow Depth (mm)",
    y = "SNODAS Snow Depth (mm)",
    title = "SNOTEL and SNODAS Daily Snow depth Relationship"
  ) +
  theme_classic()

# correlations by water year
  ggplot(wc_snodas_snotel_wint, aes(x = snotel_depth_mm, y = snodas_snow_depth_mm)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_wrap(~waterYear) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL Snow Depth (mm)",
        y = "SNODAS Snow Depth (mm)",
        title = "SNOTEL and SNODAS Daily Snow depth Relationship"
      ) +
     theme_classic()
  
```

```{r, echo = FALSE}
# Stats for snotel and snodas
snodas_snotel_stats <- wc_snodas_snotel_wint %>% 
  ungroup() %>% 
  select(-Date, -month) %>% 
  group_by(waterYear) %>% 
  mutate(rsquare_depth = 
           summary(lm(snodas_snow_depth_mm ~ snotel_depth_mm))$adj.r.squared)  %>% 
  mutate(rsquare_swe = 
            summary(lm(snodas_swe_mm ~ snotel_swe_mm))$adj.r.squared) %>% 
  mutate(rmse_depth = rmse(snodas_snow_depth_mm, snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(rmse_swe = rmse(snodas_swe_mm, snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(var_snotel_depth = var(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(var_snodas_depth = var(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(var_snotel_swe = var(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(var_snodas_swe = var(snodas_swe_mm, na.rm = TRUE)) %>% 
  mutate(med_snotel_depth = median(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(med_snodas_depth = median(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(med_snotel_swe = median(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(med_snodas_swe = median(snodas_swe_mm, na.rm = TRUE)) %>% 
  mutate(sd_snotel_depth = sd(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(sd_snodas_depth = sd(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(sd_snotel_swe = sd(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(sd_snodas_swe = sd(snodas_swe_mm, na.rm = TRUE)) %>% 
#  select(waterYear, rsquare_depth, rsquare_swe, var_depth, var_swe, rmse_depth, rmse_swe) %>% 
  slice(1)

# plot rsquared
rsquare_snodas_snotel_plot <- snodas_snotel_stats %>%
  select(waterYear, rsquare_depth, rsquare_swe) %>% 
  pivot_longer(c(rsquare_depth, rsquare_swe),
               names_to = "rsquared", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rsquared)) +
  geom_point(aes(color = rsquared)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Adjusted R2",
    title = "Adjusted R2 for SNODAS and SNOTEL by Water Year",
    color = "R2"
  ) +
  theme_classic()
rsquare_snodas_snotel_plot

# plot rmse
rmse_snodas_snotel_plot <- snodas_snotel_stats %>%
  select(waterYear, rmse_depth, rmse_swe) %>% 
  pivot_longer(c(rmse_depth, rmse_swe),
               names_to = "rmse", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rmse)) +
  geom_point(aes(color = rmse)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 1), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Root Mean Square Error",
    title = "RMSE for SNODAS and SNOTEL by Water Year",
    color = "RMSE"
  ) +
  theme_classic()
rmse_snodas_snotel_plot
```

```{r, echo = FALSe}
# compartive stats for snodas vs snotel
stats_snotel <- snodas_snotel_stats %>% 
  select(waterYear, contains("snotel")) %>% 
  select(-snotel_depth_mm, -snotel_swe_mm) %>% 
  rename(variance_depth = var_snotel_depth,
         variance_swe = var_snotel_swe,
         median_depth = med_snotel_depth,
         median_swe = med_snotel_swe,
         st_dev_depth = sd_snotel_depth,
         st_dev_swe = sd_snotel_swe) %>% 
  mutate(observation = "SNOTEL")

stats_snodas <- snodas_snotel_stats %>% 
  select(waterYear, contains("snodas")) %>% 
  select(-snodas_snow_depth_mm, -snodas_swe_mm) %>% 
  rename(variance_depth = var_snodas_depth,
         variance_swe = var_snodas_swe,
         median_depth = med_snodas_depth,
         median_swe = med_snodas_swe,
         st_dev_depth = sd_snodas_depth,
         st_dev_swe = sd_snodas_swe) %>% 
  mutate(observation = "SNODAS")
  

stats_dist_snodas_snotel <- rbind(stats_snotel, stats_snodas)

### SWE ###
variance_swe_plot <- ggplot(stats_dist_snodas_snotel, aes( x = waterYear, y = variance_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Variance",
    title = "Variance for Daily Winter SWE (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

sd_swe_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = st_dev_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Standard Deviation",
    title = "Standard Deviation of Daily Winter SWE (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

median_swe_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = median_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Median of Daily Winter SWE"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()
  
  
grid.arrange(median_swe_plot, sd_swe_plot, variance_swe_plot, ncol = 1)


### snow depth ###
variance_depth_plot <- ggplot(stats_dist_snodas_snotel, aes( x = waterYear, y = variance_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Variance",
    title = "Variance of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

sd_depth_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = st_dev_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Standard Deviation",
    title = "Standard Deviation of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

median_depth_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = median_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Median of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()
  
  
grid.arrange(median_depth_plot, sd_depth_plot, variance_depth_plot, ncol = 1)
```





Stats

```{r}
# correlation
cor.test(wc_snodas_snotel_wint$snodas_snow_depth_mm,
         wc_snodas_snotel_wint$snotel_depth_mm)

# linear model 
summary(lm(wc_snodas_snotel_wint$snodas_snow_depth_mm ~
             wc_snodas_snotel_wint$snotel_depth_mm))

# root mean square error
rmse(wc_snodas_snotel_wint$snodas_snow_depth_mm, 
     wc_snodas_snotel_wint$snotel_depth_mm, na.rm = TRUE)
```

Timing of max swe


```{r, echo = FALSE}
max_swe <- wc_snodas_snotel %>% 
  addWaterYear() 
  
# snodas
max_swe_snodas_time <- max_swe %>% 
  select(Date, waterYear, snodas_swe_mm) %>% 
  group_by(waterYear) %>% 
  top_n(1, snodas_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snodas_swe_mm) %>% 
  mutate(type = "snodas")

# SNOTEL
max_swe_snotel_time <- max_swe %>% 
  select(Date, waterYear, snotel_swe_mm) %>% 
  filter(waterYear >= 2010) %>% 
  group_by(waterYear) %>% 
  top_n(1, snotel_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snotel_swe_mm) %>% 
  mutate(type = "snotel")

max_swe_all <- rbind(max_swe_snodas_time, max_swe_snotel_time)

# plot
ggplot(max_swe_all, aes(x = waterYear, y= water_week)) +
  geom_point(aes(color = type)) +
  geom_line(aes(color = type)) +
  labs(
    title = "Timing of Peak SWE by Water Week for Whiskey Creek Station",
    x = "Water Year",
    y = "Week of Water Year"
  ) +
  theme_bw()

```



### Monthly max from Nov-Apr

```{r, echo = FALSE}
wc_snodas_snotel_monthly <- wc_snodas_snotel_wint %>% 
  group_by(waterYear, month) %>% 
  summarize(snodas_snow_depth_mm = max(snodas_snow_depth_mm),
            snodas_swe_mm = max(snodas_swe_mm),
            snotel_depth_mm = max(snotel_depth_mm),
            snotel_swe_mm = max(snotel_swe_mm)) 
```

```{r, echo = FALSE}
# correlations
ggplot(wc_snodas_snotel_monthly, aes(x = snodas_swe_mm, y = snotel_swe_mm)) +
  geom_point(aes(color = waterYear)) +
  geom_smooth(method = lm) +
  labs(
    title = "SNOTEL and SNODAS Maximum Monthly SWE Relationship"
  ) +
  theme_classic()


```

Stats

```{r}
# correlation
cor.test(wc_snodas_snotel_monthly$snodas_swe_mm,
         wc_snodas_snotel_monthly$snotel_swe_mm)

# linear model 
summary(lm(wc_snodas_snotel_monthly$snodas_swe_mm ~ wc_snodas_snotel_monthly$snotel_swe_mm))

# root mean square error
rmse(wc_snodas_snotel_monthly$snodas_swe_mm,
     wc_snodas_snotel_monthly$snotel_swe_mm, na.rm = TRUE)
```

```{r, echo = FALSE}
# correlations
ggplot(wc_snodas_snotel_monthly, aes(x = snodas_snow_depth_mm, y = snotel_depth_mm)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(
    title = "SNOTEL and SNODAS Monthly maximum snow depth Relationship"
  ) +
  theme_classic()


```

#### Comparison with PRISM data

```{r, echo = FALSE}
# read in whiskey creek prism  precip data
wc_prism_precip <- read_csv("../Data/wc_prism_precip.csv")

# whiskey creek prism tmean data
wc_prism_tmean <- read_csv("../Data/wc_prism_tmean.csv")

# combine precip adn temp
wc_prism_precip_tmean <- left_join(wc_prism_precip, wc_prism_tmean, by = "Date") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
# show how much total precipitation this area gets every month over the years
wc_prism_precip_months <- wc_prism_precip %>% 
  mutate(year = year(Date), month = month(Date)) %>% 
  group_by(year, month) %>% 
  summarize(
    monthly_precip_mm = sum(ppt)
  )

ggplot(wc_prism_precip_months, aes(x = year, y = monthly_precip_mm)) +
  geom_col() +
  facet_wrap(~month) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Whiskey Creek Total PRISM Precipitation By Month"
  ) +
  theme_bw()
```


#### freezing days
```{r, echo = FALSE}
# determine when snow and when rain
wc_prism_freeze <- wc_prism_precip_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  mutate(precip_type = ifelse(tmean <= 0, "snow", "rain")) %>% 
  filter(precip_type != is.na(precip_type)) %>% 
  mutate(prism_snow_mm = ifelse(precip_type == "snow", ppt, 0))


```

```{r, echo = FALSE}
# how much total snow vs rain per month
wc_prism_snow <- wc_prism_freeze %>% 
  filter(waterYear != 1981) %>% 
  mutate(month = month(Date)) %>% 
  group_by(waterYear, precip_type) %>% 
  summarize(
    precip_mm = sum(ppt)
  )

# how much rain vs snow
ggplot(wc_prism_snow, aes(x = waterYear, y = precip_mm)) +
  geom_point(aes(color = precip_type))+
  geom_line(aes(color = precip_type)) +
  geom_smooth(method = "lm", se = FALSE, aes(color = precip_type)) +
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    color = "Type of Precipitation",
    title = "Total PRISM Precipitation Per Water Year Whiskey Creek "
  ) +
  theme_classic()
```

```{r, echo = FALSE}
# snotel data is accumulated precipitation, so we need to sum prism snowfall to estimate how much snow
# combine with snotel data & make precipitaitona ccumulation columsn for prism data
wc_prism_snotel <- wc_clean_metric %>% 
  select(Date, "snotel_ppt_mm" = precip_mm, 
         "snotel_snow_depth_mm" = snow_depth_mm, "snotel_swe_mm" = swe_mm) %>% 
  left_join(wc_prism_freeze, by = "Date")  %>% 
  select(-waterYear) %>% 
  addWaterYear() %>% 
  mutate(month = month(Date)) %>% 
  drop_na(ppt) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(5,6,7,8,9), 0,  prism_snow_mm_accum))
 
# look at correlations for winter months
wc_prism_snotel_wint <- wc_prism_snotel %>% 
  filter(month %in% c(11,12,1,2,3,4)) 

# graph estimated accumulated prism snow with snotel swe
ggplot(wc_prism_snotel, aes(x = Date, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism")) +
  labs(
    y = "milimeters",
    title = "PRISM Estimated Snow Accumulation and SNOTEL SWE"
  ) +
  scale_x_date(date_breaks = "2 years", expand = c(0,0)) +
  theme_classic()

ggplot(wc_prism_snotel_wint, aes(x = Date, y = prism_snow_mm_accum)) +
  geom_line()+
  labs(
    y = "milimeters",
    title = "PRISM Estimated Snow Accumulation and SNOTEL SWE"
  ) +
  scale_x_date(date_labels =" %b", date_breaks = "3 months")+
  theme_classic()

# root mean square error
rmse(wc_prism_snotel_wint$prism_snow_mm_accum, wc_prism_snotel_wint$snotel_swe_mm)
```

```{r, echo = FALSE}
# plot it
wc_prism_snotel_wint_plot <- wc_prism_snotel_wint %>% 
  mutate(month_name = month(Date, label = TRUE, abbr=TRUE)) %>% 
  ggplot(aes(x = snotel_swe_mm, y = prism_snow_mm_accum)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    x = "SNOTEL SWE (mm)",
    y = "PRISM Estimated Snow Accumulation (mm)",
    title = "Whiskey Creek Station"
  ) +
  theme_bw()
wc_prism_snotel_wint_plot

# r squared
print("Adjusted R2 value is")
summary(lm(wc_prism_snotel_wint$prism_snow_mm_accum ~ wc_prism_snotel_wint$snotel_swe_mm))$adj.r.squared

# correlations by each month
wc_prism_snotel_wint_plot + facet_wrap(~factor(month_name,
                                               level = c("Nov", "Dec", "Jan", "Feb", "Mar","Apr")))

# rsquared for each month
wc_prism_snotel_wint_r2 <-  wc_prism_snotel_wint %>% 
  ungroup() %>% 
  group_by(month) %>% 
  mutate(prism_snow_snotel_swe_r2 = summary(lm(prism_snow_mm_accum ~
                                                 snotel_swe_mm))$adj.r.squared) %>% 
  slice(1) %>% 
  mutate(month_name = month(Date, label = TRUE, abbr=TRUE)) %>% 
  select(month_name, prism_snow_snotel_swe_r2)

# plot r squared values by month
ggplot(wc_prism_snotel_wint_r2, 
       aes(x = factor(month_name, level = c("Nov", "Dec", "Jan", "Feb", "Mar","Apr")), 
           y = prism_snow_snotel_swe_r2)) +
  geom_col() +
  labs(
    x = "Month",
    y = "Adjusted R2 Value",
    title = "R2 values for PRISM Estimated Snowfall and SNOTEL SWE by month"
  ) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()
```


```{r, echo = FALSE}
# max comparison
wc_prism_snotel_max <- wc_prism_snotel %>% 
  filter(waterYear != 2009) %>% 
  summarize(
    snotel_swe_mm = max(snotel_swe_mm, na.rm = TRUE),
    prism_snow_mm_accum = max(prism_snow_mm_accum)
  )

# graph estimated accumulated prism snow with snotel swe
ggplot(wc_prism_snotel_max, aes(x = waterYear, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel_swe_mm")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism_snow_mm")) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "milimeters",
    colour = "Type",
    title = "PRISM Estimated Max Accumulated Snowfall and SNOTEL Max SWE"
  ) +
  theme_classic()

# r squared
print("Adjusted R2 value is")
summary(lm(wc_prism_snotel_max$prism_snow_mm_accum ~ wc_prism_snotel_max$snotel_swe_mm))$adj.r.squared
```


#### precipitation comparison 

```{r, echo = FALSE, eval = FALSE}

ggplot(wc_prism_snotel, aes(x = snotel_ppt_mm, y = prism_ppt_mm)) +
  geom_point() +
  labs(
    x = "Snotel precipitation (mm)",
    y = "PRISM precipitation (mm)"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

```


### Beaver Spring

* Station ID:	1143

* State:	Arizona

* Network:	SNOTEL

* County:	Apache

* Elevation:	9200 ft.

* Latitude:	36.33

* Longitude:	-109.06

* HUC:	140802040401

```{r, echo = FALSE}
# grab snotel data from NRCS
beaver_spring_snotel <- grabNRCS.data(network = "SNTL",
                                      site_id = 1143,
                                      timescale = "daily",
                                      DayBgn = "2009-10-01",
                                      DayEnd = "2020-06-01")

# clean dataset
bs_clean <- beaver_spring_snotel %>% 
  select(Date,
          "air_temp_av_F" = Air.Temperature.Average..degF., 
         "air_temp_max_F" = Air.Temperature.Maximum..degF., 
        "air_temp_min_F" =  Air.Temperature.Minimum..degF.,
        "precip_in" = Precipitation.Accumulation..in..Start.of.Day.Values,
        "snow_depth_in" = Snow.Depth..in..Start.of.Day.Values,
        "swe_in" = Snow.Water.Equivalent..in..Start.of.Day.Values) %>% 
  mutate(Date = ymd(Date))

# make everything metric
# convert everything to metric units
bs_clean_metric <- bs_clean %>% 
  mutate(air_temp_av_C = round(conv_unit(air_temp_av_F, "F", "C"),2)) %>% 
  mutate(air_temp_max_C = round(conv_unit(air_temp_max_F, "F", "C"),2)) %>% 
  mutate(air_temp_min_C = round(conv_unit(air_temp_min_F, "F", "C"),2)) %>% 
  mutate(precip_mm = conv_unit(precip_in, "inch", "mm")) %>% 
  mutate(snow_depth_mm = conv_unit(snow_depth_in, "inch", "mm")) %>%
  mutate(swe_mm = conv_unit(swe_in, "inch", "mm")) %>% 
  select(Date, air_temp_av_C, air_temp_max_C, air_temp_min_C, precip_mm, snow_depth_mm, swe_mm) %>% 
  mutate(station = "beaver_springs")
```


### First explore NAs
```{r, echo = FALSE}
# table summary of missing variables
miss_var_summary(bs_clean_metric)

# plot of missing variables
vis_miss(bs_clean_metric, sort = TRUE)

# visualize NA intersections across variables with an upset plot
gg_miss_upset(bs_clean_metric, 
              nsets = 7)
```

##### Daily basic trends

```{r, echo = FALSE}
# SWE
bs_swe_daily <- ggplot(bs_clean, aes(x = Date, y = swe_in)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (in)",
                      title = "Daily SWE"
                    ) +
                    theme_classic()

# snow depth
bs_depth_daily <- ggplot(bs_clean, aes(x = Date, y = snow_depth_in)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (in)",
                      title = "Daily Snow Depth"
                    ) +
                    theme_classic()

# max temp
bs_max_temp_daily <- ggplot(bs_clean, aes(x = Date, y = air_temp_max_F)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (F)",
                      title = "Daily Maximum Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(bs_swe_daily, bs_depth_daily, bs_max_temp_daily, ncol = 2)

```



#### Monthly basic trends

```{r, echo = FALSE}
# maximum monthly data
bs_max_monthly <- bs_clean_metric %>% 
  group_by(year = year(Date), 
           month = month(Date)) %>% 
  summarize(air_temp_av_C = max(air_temp_av_C, na.rm = TRUE),
            air_temp_max_C = max(air_temp_max_C, na.rm = TRUE),
            air_temp_min_C = max(air_temp_min_C, na.rm = TRUE),
            swe_mm = max(swe_mm, na.rm = TRUE),
            snow_depth_mm = max(snow_depth_mm, na.rm = TRUE)) %>% 
  mutate(Date = ymd(paste0(year, "-", month, "-01")))

# SWE
bs_swe_monthly <- ggplot(bs_max_monthly, aes(x = Date, y = swe_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (mm)",
                      title = "Max Monthly SWE"
                    ) +
                    theme_classic()

# snow depth
bs_depth_monthly <- ggplot(bs_max_monthly, aes(x = Date, y = snow_depth_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (mm)",
                      title = "Max Monthly Snow Depth"
                    ) +
                    theme_classic()

# max temp
bs_max_temp_monthly <- ggplot(bs_max_monthly,
                              aes(x = Date, y = air_temp_max_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Maximum Air Temperature"
                    ) +
                    theme_classic()


# max average temp
bs_max_av_temp_monthly <- ggplot(bs_max_monthly,
                              aes(x = Date, y = air_temp_av_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Average Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(bs_swe_monthly, bs_depth_monthly, bs_max_temp_monthly, bs_max_av_temp_monthly, ncol = 2)


```

## Whiskey creek vs Beaver springs
```{r, echo = FALSE}
# stack the two snotel station data
wc_bs <- wc_clean_metric %>% 
  mutate(station = "whiskey_creek") %>% 
  rbind(bs_clean_metric) %>% 
  addWaterYear() %>% 
  filter(waterYear > 2009)

# SWE
wc_bs_swe_plot <- ggplot(wc_bs, aes(x = Date, y = swe_mm)) +
                geom_line(aes(color = station)) +
                labs(
                  y = "SWE (mm)", 
                  x = "Date",
                  color = "Station"
                ) +
                scale_y_continuous(expand = c(0,0)) +
                scale_x_date(expand = c(0,0), date_breaks = "2 years")+
                theme_bw()

# Snow depth
wc_bs_depth_plot <-ggplot(wc_bs, aes(x = Date, y = snow_depth_mm)) +
                  geom_line(aes(color = station)) +
                  labs(
                    y = "Snow Depth (mm)", 
                    x = "Date",
                    color = "Station"
                  ) +
                  scale_y_continuous(expand = c(0,0)) +
                  scale_x_date(expand = c(0,0), date_breaks = "2 years")+
                  theme_bw()


grid.arrange(wc_bs_swe_plot, wc_bs_depth_plot, ncol = 1)
```

#### Seasonal trends for both snotel sites
```{r, echo = FALSE}
# clean for seasonal calculations
wc_bs_swe_seasonal <- wc_bs %>% 
  select(Date, waterYear, swe_mm, snow_depth_mm, station) %>% 
  mutate(month = month(Date)) %>% 
  mutate(day = day(Date)) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  filter(month %in% c(1,2,3,4,11,12)) 

# daily average swe
wc_bs_swe_seasonal_daily <- wc_bs_swe_seasonal %>% 
            group_by(station, month, day) %>% 
            summarize(
              mean_swe = mean(swe_mm, na.rm = TRUE),
              mean_depth = mean(snow_depth_mm, na.rm = TRUE)
            ) %>% 
            mutate(fake_md = paste0(month,"-", day)) %>% 
            mutate(fake_year = ifelse(month %in% c(11,12), 1999, 2000)) %>% 
            mutate(fake_date = mdy(paste0(fake_md, "-", fake_year)))

# average daily swe from 2009-2020
ggplot(wc_bs_swe_seasonal_daily, aes(x = fake_date, y = mean_swe)) +
  geom_line(aes(color = station)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  labs(
    x = "Month",
    y = "Mean SWE (mm)"
  )

# weekly average SWE
wc_bs_swe_seasonal_week <- wc_bs_swe_seasonal %>% 
  group_by(station, water_week) %>% 
  summarize(
    mean_swe = mean(swe_mm, na.rm = TRUE),
    mean_depth = mean(snow_depth_mm, na.rm = TRUE)
  )

# average weekly swe from 2009-2020
ggplot(wc_bs_swe_seasonal_week, aes(x = water_week, y = mean_swe)) +
  geom_line(aes(color = station)) +
  labs(
    x = "Week of Water Year",
    y = "Mean SWE (mm)"
  ) +
  theme_classic()
```


```{r, echo = FALSE}
# look at the correlation scatter plot between the swe and depth
wc_bs_swe_depth <- wc_clean_metric %>% 
  select(Date, "swe_mm_wc" = swe_mm, "snow_depth_mm_wc" = snow_depth_mm) %>% 
  left_join(bs_clean_metric, by = "Date") %>% 
  select(Date, swe_mm_wc, snow_depth_mm_wc, "swe_mm_bs" = swe_mm, "snow_depth_mm_bs" = snow_depth_mm)

 wc_bs_swe_corr <- ggplot(wc_bs_swe_depth, aes( x = swe_mm_wc, y = swe_mm_bs)) +
          geom_point() +
          geom_smooth(method= "lm", se = FALSE) +
          labs(
            x = "Whiskey Creek SWE (mm)",
            y = "Beaver Springs SWE (mm)"
          ) +
          theme_bw()
 
wc_bs_depth_corr <- ggplot(wc_bs_swe_depth, aes( x = snow_depth_mm_wc, y = snow_depth_mm_bs)) +
          geom_point() +
          geom_smooth(method= "lm", se = FALSE) +
          labs(
            x = "Whiskey Creek Snow Depth (mm)",
            y = "Beaver Springs Snow Depth (mm)"
          ) +
          theme_bw()

grid.arrange(wc_bs_swe_corr, wc_bs_depth_corr, ncol = 1)
```



### timing of peak swe

```{r, echo = FALSE}
ws_bs_peak_swe <- wc_bs %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  ungroup() %>% 
  select(waterYear, water_week, swe_mm, station) %>% 
  group_by(station, waterYear) %>% 
  top_n(1, swe_mm) %>% 
  top_n(1, water_week)

ggplot(ws_bs_peak_swe, aes(x = waterYear, y = water_week)) +
  geom_point(aes(color = station)) +
  geom_line(aes(color = station)) +
  labs(
    x = "Water Year",
    y = "Water Week"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =2)) +
  theme_classic()

```



### SNOTEL vs SNODAS

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# read in snodas swe
bs_snodas_swe <- read_csv("../Data/beaver_springs_snodas_swe.csv",
                          col_names = c("Date", "snodas_swe_mm"), 
                          skip = 1)
# read in snodas snow depth
bs_snodas_depth <- read_csv("../Data/beaver_springs_snodas_depth.csv",
                          col_names = c("Date", "snodas_snow_depth_mm"), 
                          skip = 1)
```

#### Daily data

```{r, echo = FALSE}
# combine snodas and snotel
bs_snodas_snotel <- left_join(bs_snodas_swe, bs_snodas_depth) %>% 
  left_join(wc_clean_metric, by = "Date") %>% 
  select(Date, snodas_swe_mm,  snodas_snow_depth_mm,
         "snotel_depth_mm" = snow_depth_mm, 
         "snotel_swe_mm" = swe_mm) %>% 
  addWaterYear()

# tidy the data for plotting
bs_tidy_snodas_snotel <- bs_snodas_snotel %>% 
  pivot_longer(c(snodas_swe_mm, snodas_snow_depth_mm, 
                 snotel_depth_mm, snotel_swe_mm),
               names_to = "type", values_to = "value_mm") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
### SWE ###
bs_das_tel_swe <- bs_tidy_snodas_snotel %>% 
  filter(type == "snodas_swe_mm" | type ==  "snotel_swe_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "SWE (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL SWE Beaver Springs Station"
  ) +
  theme_classic()
wc_das_tel_swe

# clean the data for graphing
bs_snodas_snotel_wint <- bs_snodas_snotel %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  filter(waterYear > 2009)# 2009 not a full water year
  
# all correlation    
  ggplot(bs_snodas_snotel_wint, aes(x = snotel_swe_mm, y = snodas_swe_mm)) +
    geom_point(aes(color = waterYear)) +
    geom_smooth(method = lm) +
    scale_x_continuous(expand = c(0,0)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(
      x = "SNOTEL SWE (mm)",
      y = "SNODAS SWE (mm)",
      title = "SNOTEL and SNODAS Daily SWE Relationship Nov-Apr"
    ) +
    theme_classic()

# correlations by water year
  ggplot(bs_snodas_snotel_wint, aes(x = snotel_swe_mm, y = snodas_swe_mm)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_wrap(~waterYear) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL SWE (mm)",
        y = "SNODAS SWE (mm)",
        title = "SNOTEL and SNODAS Daily SWE Relationship Nov-Apr"
      ) +
      theme_classic()
```

```{r, echo = FALSE}
# snow depth
bs_das_tel_depth <- bs_tidy_snodas_snotel %>% 
  filter(type == "snodas_snow_depth_mm" | type ==  "snotel_depth_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "Snow depth (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL Snow Depth"
  ) +
  theme_classic()
bs_das_tel_depth

# correlation
ggplot(bs_snodas_snotel_wint,
       aes(x = snotel_depth_mm, y = snodas_snow_depth_mm)) +
  geom_point(aes(color = waterYear)) +
  geom_smooth(method = lm) +
  labs(
    x = "SNOTEL Snow Depth (mm)",
    y = "SNODAS Snow Depth (mm)",
    title = "SNOTEL and SNODAS Daily Snow depth Relationship"
  ) +
  theme_classic()

# correlations by water year
  ggplot(bs_snodas_snotel_wint, aes(x = snotel_depth_mm, 
                                    y = snodas_snow_depth_mm)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_wrap(~waterYear) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL Snow Depth (mm)",
        y = "SNODAS Snow Depth (mm)",
        title = "SNOTEL and SNODAS Daily Snow depth Relationship"
      ) +
     theme_classic()
```


```{r, echo = FALSE}
# r squared for each water year for snodas vs snotel swe and depth
rsquare_snodas_snotel_bs <- bs_snodas_snotel_wint %>% 
  ungroup() %>% 
  select(-Date, -month) %>% 
  group_by(waterYear) %>% 
  mutate(rsquare_depth = 
           summary(lm(snodas_snow_depth_mm ~ snotel_depth_mm))$adj.r.squared)  %>% 
  mutate(rsquare_swe = 
            summary(lm(snodas_swe_mm ~ snotel_swe_mm))$adj.r.squared) %>% 
  select(waterYear, rsquare_depth, rsquare_swe) %>% 
  slice(1)

# plot rsquared
rsquare_snodas_snotel_plot_bs <- rsquare_snodas_snotel_bs %>%
  pivot_longer(c(rsquare_depth, rsquare_swe),
               names_to = "rsquared", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rsquared)) +
  geom_point(aes(color = rsquared)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Adjusted R2",
    title = "Adjusted R2 for SNODAS and SNOTEL by Water Year",
    color = "R2"
  ) +
  theme_classic()
rsquare_snodas_snotel_plot_bs


```


```{r, echo = FALSE}
max_swe <- bs_snodas_snotel %>% 
  addWaterYear() 
  
# snodas
max_swe_snodas_time <- max_swe %>% 
  select(Date, waterYear, snodas_swe_mm) %>% 
  group_by(waterYear) %>% 
  top_n(1, snodas_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snodas_swe_mm) %>% 
  mutate(type = "snodas")

# SNOTEL
max_swe_snotel_time <- max_swe %>% 
  select(Date, waterYear, snotel_swe_mm) %>% 
  filter(waterYear >= 2010) %>% 
  group_by(waterYear) %>% 
  top_n(1, snotel_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snotel_swe_mm) %>% 
  mutate(type = "snotel")

max_swe_all <- rbind(max_swe_snodas_time, max_swe_snotel_time)

# plot
ggplot(max_swe_all, aes(x = waterYear, y= water_week)) +
  geom_point(aes(color = type)) +
  geom_line(aes(color = type)) +
  labs(
    title = "Timing of Peak SWE by Water Week for Beaver Springs Station",
    x = "Water Year",
    y = "Week of Water Year"
  ) +
  theme_bw()

```

### Comparison with PRISM data

```{r, echo = FALSE}
# read in beaver springs prism precipiation data
bs_prism_precip <- read_csv("../Data/bs_prism_precip.csv")

# read in beaver springs prism temperature data
bs_prism_tmean <- read_csv("../Data/bs_prism_tmean.csv")

# combine precip adn temp
bs_prism_precip_tmean <- left_join(bs_prism_precip, bs_prism_tmean, by = "Date") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
# show how much total precipitation this area gets every month over the years
bs_prism_precip_months <- bs_prism_precip %>% 
  mutate(year = year(Date), month = month(Date)) %>% 
  group_by(year, month) %>% 
  summarize(
    monthly_precip_mm = sum(ppt)
  )

ggplot(bs_prism_precip_months, aes(x = year, y = monthly_precip_mm)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE)+
  facet_wrap(~month) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Beaver Springs Total PRISM Precipitation By Month"
  ) +
  theme_bw()
```

#### freezing days
```{r, echo = FALSE}

bs_prism_freeze <- bs_prism_precip_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  mutate(precip_type = ifelse(tmean <= 0, "snow", "rain")) %>% 
  filter(precip_type != is.na(precip_type)) %>% 
  mutate(prism_snow_mm = ifelse(precip_type == "snow", ppt, 0))

# how much total snow vs rain per month
bs_prism_snow <- bs_prism_freeze %>% 
  filter(waterYear != 1981) %>% 
  mutate(month = month(Date)) %>% 
  group_by(waterYear, precip_type) %>% 
  summarize(
    precip_mm = sum(ppt)
  )

# how much rain vs snow
ggplot(bs_prism_snow, aes(x = waterYear, y = precip_mm)) +
  geom_point(aes(color = precip_type))+
  geom_line(aes(color = precip_type)) +
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    color = "Type of Precipitation",
    title = "Total PRISM Precipitation Per Water Year Beaver Springs"
  ) +
  theme_classic()
```

#### PRISM Snow accumulation vs SNOTEL SWE

```{r, echo = FALSE}
# snotel data is accumulated precipitation, so we need to sum prism snowfall to estimate how much snow
# combine with snotel data & make precipitaitona ccumulation columsn for prism data
bs_prism_snotel <- bs_clean_metric %>% 
  select(Date, "snotel_ppt_mm" = precip_mm, 
         "snotel_snow_depth_mm" = snow_depth_mm, "snotel_swe_mm" = swe_mm) %>% 
  left_join(bs_prism_freeze, by = "Date")  %>% 
  select(-waterYear) %>% 
  addWaterYear() %>% 
  mutate(month = month(Date)) %>% 
  drop_na(ppt) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm))  # accumulate prism precip
 
# look at correlations for winter months
bs_prism_snotel_wint <- bs_prism_snotel %>% 
  filter(month %in% c(11,12,1,2,3,4)) 

# graph estimated accumulated prism snow with snotel swe
ggplot(bs_prism_snotel_wint, aes(x = Date, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism")) +
  labs(
    y = "milimeters",
    title = "PRISM Estimated Snow Accumulation and SNOTEL SWE"
  ) +
  theme_classic()

# root mean square error
rmse(bs_prism_snotel_wint$prism_snow_mm_accum, bs_prism_snotel_wint$snotel_swe_mm)

# plot it
bs_prism_snotel_wint_plot <- bs_prism_snotel_wint %>% 
  ggplot(aes(x = snotel_swe_mm, y = prism_snow_mm_accum)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    x = "SNOTEL SWE (mm)",
    y = "PRISM Estimated Snow Accumulation (mm)",
    title = "Beaver Springs Station"
  ) +
  theme_bw()
bs_prism_snotel_wint_plot

# r squared
print("Adjusted R2 value is")
summary(lm(bs_prism_snotel_wint$prism_snow_mm_accum ~ bs_prism_snotel_wint$snotel_swe_mm))$adj.r.squared

# correlations by each month
bs_prism_snotel_wint_plot + facet_wrap(~month) 

# rsquared for each month
bs_prism_snotel_wint_r2 <-  bs_prism_snotel_wint %>% 
  ungroup() %>% 
  group_by(month) %>% 
  mutate(prism_snow_snotel_swe_r2 = summary(lm(prism_snow_mm_accum ~
                                                 snotel_swe_mm))$adj.r.squared) %>% 
  slice(1) %>% 
  mutate(month_name = month(Date, label = TRUE, abbr=TRUE)) %>% 
  select(month_name, prism_snow_snotel_swe_r2)

# plot r squared values by month
ggplot(bs_prism_snotel_wint_r2, 
       aes(x = factor(month_name, level = c("Nov", "Dec", "Jan", "Feb", "Mar","Apr")), 
           y = prism_snow_snotel_swe_r2)) +
  geom_col() +
  labs(
    x = "Month",
    y = "Adjusted R2 Value",
    title = "R2 values for PRISM Estimated Snowfall and SNOTEL SWE by month"
  ) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()
```

```{r, echo = FALSE}
# max comparison
bs_prism_snotel_max <- bs_prism_snotel %>% 
  filter(waterYear != 2009) %>% 
  summarize(
    snotel_swe_mm = max(snotel_swe_mm, na.rm = TRUE),
    prism_snow_mm_accum = max(prism_snow_mm_accum)
  )

# graph estimated accumulated prism snow with snotel swe
ggplot(bs_prism_snotel_max, aes(x = waterYear, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel_swe_mm")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism_snow_mm")) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "milimeters",
    colour = "Type",
    title = "PRISM Estimated Max Accumulated Snowfall and SNOTEL Max SWE"
  ) +
  theme_classic()

# r squared
print("Adjusted R2 value is")
summary(lm(bs_prism_snotel_max$prism_snow_mm_accum ~ bs_prism_snotel_max$snotel_swe_mm))$adj.r.squared
```

#### precipitation comparison 

```{r, eval = FALSE}

# combine with snotel data
bs_prism_snotel <- bs_prism_precip %>% 
  mutate(Date = ymd(Date)) %>% 
  rename(prism_ppt_mm = ppt) %>% 
  left_join(bs_clean_metric, by = "Date") %>% 
  select(Date, prism_ppt_mm, "snotel_ppt_mm" = precip_mm, snow_depth_mm, swe_mm)


```