---
title: "Snotel"
author: "AnnaClaire Marley"
date: "6/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This RMarkdown conducts initial time series analyses on two SNOTEL sites in the Chuska Mountains located in the Navajo Nation. It then compares SNOTEL data to remote sensing data such as SNODAS and PRISM to understand which products best account for snow.

**Read in packages**

```{r, message=FALSE, include = FALSE}
library(RNRCS)
library(lubridate)
library(gridExtra)
library(measurements)
library(mltools)
library(dataRetrieval)
library(naniar)
library(ggridges)
library(tidyverse)
library(janitor)
library(kableExtra)
library(Metrics)
```

```{r, include = FALSE}
# set path to source funcitons from data cleaning folder
path_clean = "../Functions/Cleaning/"
path_graph = "../Functions/Graphing/"

mltools::rmse
```

```{r, echo = FALSE}
# source functions
source(paste0(path_clean, "to_water_week.R")) # for calculating week of water year
source(paste0(path_graph, "graph_season_trend.R"))
source(paste0(path_graph, "graph_correlation.R"))
```

```{r, echo = FALSE}
# for organizing graphs in month order of water years
wy_order = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")

# colors for months
colors = c('#F6BD60', '#7FB069', '#548687', "plum4", "coral2", "slategrey")

```


### Navajo Whiskey Creek

* Station ID:	1138

* State:	New Mexico

* Network:	SNOTEL

* County:	San Juan

* Elevation:	9050 ft.

* Latitude:	36.18

* Longitude:	-108.95

* HUC:	140802040206

```{r, echo = FALSE}
# fixes numbers showing up in scientific notation
options(scipen = 999)

# grab snotel data from NRCS
whiskey_creek_snotel <- grabNRCS.data(network = "SNTL",
                                      site_id = 1138,
                                      timescale = "daily",
                                      DayBgn = "1984-10-31",
                                      DayEnd = "2020-06-01")

# clean dataset
wc_clean <- whiskey_creek_snotel %>% 
  select(Date,
          "air_temp_av_F" = Air.Temperature.Average..degF., 
         "air_temp_max_F" = Air.Temperature.Maximum..degF., 
        "air_temp_min_F" =  Air.Temperature.Minimum..degF.,
        "precip_in" = Precipitation.Accumulation..in..Start.of.Day.Values,
        "snow_depth_in" = Snow.Depth..in..Start.of.Day.Values,
        "swe_in" = Snow.Water.Equivalent..in..Start.of.Day.Values) %>% 
  mutate(Date = ymd(Date)) %>% 
  mutate(snow_depth_in = ifelse(snow_depth_in < 100, 
                                snow_depth_in, NA)) # get rid of outliers

# convert everything to metric units
wc_clean_metric <- wc_clean %>% 
  mutate(air_temp_av_C = round(conv_unit(air_temp_av_F, "F", "C"),2)) %>% 
  mutate(air_temp_max_C = round(conv_unit(air_temp_max_F, "F", "C"),2)) %>% 
  mutate(air_temp_min_C = round(conv_unit(air_temp_min_F, "F", "C"),2)) %>% 
  mutate(precip_mm = conv_unit(precip_in, "inch", "mm")) %>% 
  mutate(snow_depth_mm = conv_unit(snow_depth_in, "inch", "mm")) %>%
  mutate(swe_mm = conv_unit(swe_in, "inch", "mm")) %>% 
  select(Date, air_temp_av_C, air_temp_max_C, air_temp_min_C, precip_mm, snow_depth_mm, swe_mm) 
```

### First explore NAs
```{r, echo = FALSE}
# table summary of missing variables
miss_var_summary(wc_clean_metric)

# plot of missing variables
vis_miss(wc_clean_metric, sort = TRUE)

# visualize NA intersections across variables with an upset plot
gg_miss_upset(wc_clean_metric, 
              nsets = 6)
```

- The horizontal black bars in the lower left indicate the number of NAs total for each variable, with the variable name shown to the right of the corresponding horizontal bar

- The vertical black lines with black dots indicate the variables between which the frequency of co-occurring NAs is indicated by the vertical black bars


#### Daily basic trends

```{r, echo = FALSE}
# SWE
wc_swe_daily <- ggplot(wc_clean_metric, aes(x = Date, y = swe_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (mm)",
                      title = "Daily SWE"
                    ) +
                    theme_classic()

# snow depth
wc_depth_daily <- ggplot(wc_clean_metric, aes(x = Date, y = snow_depth_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (mm)",
                      title = "Daily Snow Depth"
                    ) +
                    theme_classic()

# max temp
wc_max_temp_daily <- ggplot(wc_clean_metric, aes(x = Date, y = air_temp_max_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Daily Maximum Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(wc_swe_daily, wc_depth_daily, wc_max_temp_daily, ncol = 2)

```

#### ENSO
```{r, echo = FALSE}

el_nino_yrs <- c(2010, 2015, 2016, 2019, 2020)
la_nina_yrs <- c(2008, 2009, 2011, 2012, 2017, 2018)

wc_enso <- wc_clean_metric %>% 
  addWaterYear() %>% 
  mutate(ENSO = ifelse(waterYear %in% c(el_nino_yrs), "el_nino", 
                       ifelse(waterYear %in% c(la_nina_yrs), 'la_nina', NA)))

# make rectangles over swe plot to show when el nino and la nina
ggplot() +
  geom_line(data = wc_enso, aes(x = Date, y = swe_mm)) +
  # el nino years
  geom_rect(aes(xmin = as.Date('2009-10-01'), ymin = -Inf, 
                 xmax = as.Date('2010-09-30'), ymax = Inf),
             alpha = 0.3, fill = "tomato1") +
  geom_rect(aes(xmin = as.Date('2014-10-01'), ymin = -Inf, 
                 xmax = as.Date('2015-09-30'), ymax = Inf),
             alpha = 0.3, fill = "tomato1") +
  geom_rect(aes(xmin = as.Date('2015-10-01'), ymin = -Inf, 
                 xmax = as.Date('2016-09-30'), ymax = Inf),
             alpha = 0.3, fill = "tomato1") +
  geom_rect(aes(xmin = as.Date('2018-10-01'), ymin = -Inf, 
                 xmax = as.Date('2019-09-30'), ymax = Inf),
             alpha = 0.3, fill = "tomato1") +
  geom_rect(aes(xmin = as.Date('2019-10-01'), ymin = -Inf, 
                 xmax = as.Date('2020-09-30'), ymax = Inf),
             alpha = 0.3, fill = "tomato1") +
  # la nina years
  geom_rect(aes(xmin = as.Date('2010-10-01'), ymin = -Inf, 
                 xmax = as.Date('2011-09-30'), ymax = Inf),
             alpha = 0.3, fill = "skyblue2") +
  geom_rect(aes(xmin = as.Date('2011-10-01'), ymin = -Inf, 
                 xmax = as.Date('2012-09-30'), ymax = Inf),
             alpha = 0.3, fill = "skyblue2") +
  geom_rect(aes(xmin = as.Date('2016-10-01'), ymin = -Inf, 
                 xmax = as.Date('2017-09-30'), ymax = Inf),
             alpha = 0.3, fill = "skyblue2") +
  geom_rect(aes(xmin = as.Date('2017-10-01'), ymin = -Inf, 
                 xmax = as.Date('2018-09-30'), ymax = Inf),
             alpha = 0.3, fill = "skyblue2") +
  labs(
    y = "SWE (mm)"
  ) +
  scale_x_date(expand = c(0,0)) +
  theme_classic()

```

blue = la nina years

red = el nino years


```{r, echo = FALSE}

# fun visualization from Alison:
wc_dist <- wc_clean_metric %>% 
  addWaterYear() %>% 
  mutate(waterYear = as.character(waterYear)) %>% 
  filter(month(Date) %in% c(11,12,1,2,3,4)) %>% 
  group_by(waterYear) %>% 
  mutate(median = median(swe_mm)) %>% 
  filter(swe_mm != 0) %>% 
  mutate(max = max(swe_mm)) %>% 
  mutate(year = year(Date))

ggplot(wc_dist, aes(x = swe_mm, y = fct_reorder(waterYear, -year))) +
  geom_density_ridges(scale = 7,
                      aes(fill = waterYear),
                      size = 0.3,
                      color = "NA",
                      alpha = 0.5,
                      show.legend = FALSE) +
 # scale_fill_manual(values = pal_4) +
 # scale_color_manual(values = pal_4) +
#  scale_x_continuous(breaks = c(0,1,2,3,4,5), limits = c(0,5), expand = c(0,0)) +
  theme_minimal() +
  labs(
    x = "SWE (mm)",
    y = "Water Years"
  )
  
  


```


#### Monthly basic trends

```{r, echo = FALSE}
# maximum monthly data
wc_max_monthly <- wc_clean_metric %>% 
  group_by(year = year(Date), 
           month = month(Date)) %>% 
  summarize(air_temp_av_C = max(air_temp_av_C, na.rm = TRUE),
            air_temp_max_C = max(air_temp_max_C, na.rm = TRUE),
            air_temp_min_C = max(air_temp_min_C, na.rm = TRUE),
            swe_mm = max(swe_mm, na.rm = TRUE),
            snow_depth_mm = max(snow_depth_mm, na.rm = TRUE)) %>% 
  mutate(Date = ymd(paste0(year, "-", month, "-01")))

# SWE
wc_swe_monthly <- ggplot(wc_max_monthly, aes(x = Date, y = swe_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (mm)",
                      title = "Max Monthly SWE"
                    ) +
                    theme_classic()

# snow depth
wc_depth_monthly <- ggplot(wc_max_monthly, aes(x = Date, y = snow_depth_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (mm)",
                      title = "Max Monthly Snow Depth"
                    ) +
                    theme_classic()

# max temp
wc_max_temp_monthly <- ggplot(wc_max_monthly,
                              aes(x = Date, y = air_temp_max_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Maximum Air Temperature"
                    ) +
                    theme_classic()


# max average temp
wc_max_av_temp_monthly <- ggplot(wc_max_monthly,
                              aes(x = Date, y = air_temp_av_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Average Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(wc_swe_monthly, wc_depth_monthly, wc_max_temp_monthly, wc_max_av_temp_monthly, ncol = 2)


```

#### Water year trends
```{r, echo = FALSE}
# water year max values
wc_max_wy <- wc_max_monthly %>% 
  addWaterYear() %>% 
  ungroup() %>% 
  group_by(waterYear) %>% 
  summarize(
    swe_mm = max(swe_mm),
    snow_depth_mm = max(snow_depth_mm)
  ) %>% 
  filter(waterYear > 2009)

ggplot(wc_max_wy, aes(x = waterYear, y = swe_mm)) +
  geom_point() +
  geom_line()
```


### Compare to SNODAS data (extracted for the same lat/long as snotel site)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# read in snodas swe
wc_snodas_swe <- read_csv("../Data/whiskey_creek_snodas_swe.csv",
                          col_names = c("Date", "snodas_swe_mm"), 
                          skip = 1)
# read in snodas snow depth
wc_snodas_depth <- read_csv("../Data/whiskey_creek_snodas_depth.csv",
                          col_names = c("Date", "snodas_snow_depth_mm"), 
                          skip = 1)
```

#### Daily data

```{r, echo = FALSE}
# combine snodas and snotel
wc_snodas_snotel <- left_join(wc_snodas_swe, wc_snodas_depth) %>% 
  left_join(wc_clean_metric, by = "Date") %>% 
  select(Date, snodas_swe_mm,  snodas_snow_depth_mm,
         "snotel_depth_mm" = snow_depth_mm, 
         "snotel_swe_mm" = swe_mm) %>% 
  addWaterYear()

# tidy the data for plotting
wc_tidy_snodas_snotel <- wc_snodas_snotel %>% 
  pivot_longer(c(snodas_swe_mm, snodas_snow_depth_mm, 
                 snotel_depth_mm, snotel_swe_mm),
               names_to = "type", values_to = "value_mm") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
# look at the distribution in SWE in snotel compared to snodas
wc_snotel_snodas_dist <- wc_tidy_snodas_snotel %>% 
  filter(waterYear > 2009) %>% 
  filter(month(Date) %in% c(11,12,1,2,3,4))

# swe distributions
wc_snotel_snodas_dist %>% 
  filter(type == "snodas_swe_mm" | type == "snotel_swe_mm") %>% 
  ggplot(., aes(x = value_mm, y = type)) +
  geom_density_ridges(scale = 7,
                      aes(fill = type),
                      size = 0.3,
                      color = "NA",
                      alpha = 0.5,
                      show.legend = FALSE) +
  theme_minimal() +
  labs(
    x = "SWE (mm)",
    y = "Station"
  )
  
# depth distribution
wc_snotel_snodas_dist %>% 
  filter(type == "snodas_snow_depth_mm" | type == "snotel_depth_mm") %>% 
  ggplot(aes(x = value_mm, y = type)) +
  geom_density_ridges(scale = 7,
                      aes(fill = type),
                      size = 0.3,
                      color = "NA",
                      alpha = 0.5,
                      show.legend = FALSE) +
  theme_minimal() +
  labs(
    x = "Snow Depth (mm)",
    y = "Station"
  )
```


```{r, echo = FALSE}
### SWE ###
wc_das_tel_swe <- wc_tidy_snodas_snotel %>% 
  filter(type == "snodas_swe_mm" | type ==  "snotel_swe_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "SWE (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL SWE Whiskey Creek Station"
  ) +
  theme_classic()
wc_das_tel_swe

# clean the data for graphing
wc_snodas_snotel_wint <- wc_snodas_snotel %>% 
  mutate(month = month(Date)) %>% 
  mutate(month_name = month(Date, label = TRUE, abbr = TRUE)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  filter(waterYear > 2009)# 2009 not a full water year

# rmse and mae by month between snodas swe and snotel swe
wc_snodas_snotel_wint_rmse_mae <- wc_snodas_snotel_wint %>% 
  ungroup() %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(snodas_swe_mm, snotel_swe_mm)) %>% 
  mutate(mae = mae(snodas_swe_mm, snotel_swe_mm)) %>% 
  select(month_name, rmse, mae) %>% 
  slice(1)

# rmse and mae by water year between snodas swe and snotel swe
wc_snodas_snotel_wint_rmse_mae_wy <- wc_snodas_snotel_wint %>% 
  ungroup() %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(snodas_swe_mm, snotel_swe_mm)) %>% 
  mutate(mae = mae(snodas_swe_mm, snotel_swe_mm)) %>% 
  select(waterYear, rmse, mae) %>% 
  slice(1)
  
# all    
all_wc_snodas <-  graph_correlation(wc_snodas_snotel_wint, x_data = wc_snodas_snotel_wint$snodas_swe_mm, 
                        y_data = wc_snodas_snotel_wint$snotel_swe_mm,
                        title = "Whiskey Creek SNOTEL and SNODAS Daily SWE", 
                         xlabel = "SNODAS SWE (mm)",
                         ylabel = "SNOTEL SWE (mm)") +
                          geom_point(aes(color = month_name)) +
                          scale_color_manual(values = colors)
 
all_wc_snodas

# by water year
all_wc_snodas + facet_wrap(~waterYear)
  
# by month  
all_wc_snodas + facet_wrap(~month_name) 
  
  # correlations for 2014 and 2018
wc_snodas_snotel_wint %>% 
  filter(waterYear %in% c(2018)) %>% 
   ggplot(aes(x = snotel_swe_mm, y = snodas_swe_mm)) +
      geom_point() +
      geom_smooth(method = lm, se = FALSE) +
      facet_wrap(~waterYear) +
   #   scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL SWE (mm)",
        y = "SNODAS SWE (mm)",
        title = "SNOTEL and SNODAS Daily SWE Relationship Nov-Apr"
      ) +
      theme_classic()
  
```

```{r, echo = FALSE}
# snow depth
wc_das_tel_depth <- wc_tidy_snodas_snotel %>% 
  filter(type == "snodas_snow_depth_mm" | type ==  "snotel_depth_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "Snow depth (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL Snow Depth"
  ) +
  theme_classic()
wc_das_tel_depth

# correlation
ggplot(wc_snodas_snotel_wint, aes(x = snotel_depth_mm, y = snodas_snow_depth_mm)) +
  geom_point(aes(color = waterYear)) +
  geom_smooth(method = lm) +
  labs(
    x = "SNOTEL Snow Depth (mm)",
    y = "SNODAS Snow Depth (mm)",
    title = "SNOTEL and SNODAS Daily Snow depth Relationship"
  ) +
  theme_classic()

# correlations by water year
  ggplot(wc_snodas_snotel_wint, aes(x = snotel_depth_mm, y = snodas_snow_depth_mm)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_wrap(~waterYear) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL Snow Depth (mm)",
        y = "SNODAS Snow Depth (mm)",
        title = "SNOTEL and SNODAS Daily Snow depth Relationship"
      ) +
     theme_classic()
  
```



```{r, echo = FALSE}
# Stats for snotel and snodas
snodas_snotel_stats <- wc_snodas_snotel_wint %>% 
  ungroup() %>% 
  select(-Date, -month) %>% 
  group_by(waterYear) %>% 
  mutate(rsquare_depth = 
           summary(lm(snodas_snow_depth_mm ~ snotel_depth_mm))$adj.r.squared)  %>% 
  mutate(rsquare_swe = 
            summary(lm(snodas_swe_mm ~ snotel_swe_mm))$adj.r.squared) %>% 
  mutate(rmse_depth = mltools::rmse(snodas_snow_depth_mm, snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(rmse_swe = mltools::rmse(snodas_swe_mm, snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(var_snotel_depth = var(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(var_snodas_depth = var(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(var_snotel_swe = var(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(var_snodas_swe = var(snodas_swe_mm, na.rm = TRUE)) %>% 
  mutate(med_snotel_depth = median(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(med_snodas_depth = median(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(med_snotel_swe = median(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(med_snodas_swe = median(snodas_swe_mm, na.rm = TRUE)) %>% 
  mutate(sd_snotel_depth = sd(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(sd_snodas_depth = sd(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(sd_snotel_swe = sd(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(sd_snodas_swe = sd(snodas_swe_mm, na.rm = TRUE)) %>% 
#  select(waterYear, rsquare_depth, rsquare_swe, var_depth, var_swe, rmse_depth, rmse_swe) %>% 
  slice(1)

# plot rsquared
rsquare_snodas_snotel_plot <- snodas_snotel_stats %>%
  select(waterYear, rsquare_depth, rsquare_swe) %>% 
  pivot_longer(c(rsquare_depth, rsquare_swe),
               names_to = "rsquared", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rsquared)) +
  geom_point(aes(color = rsquared)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Adjusted R2",
    title = "Adjusted R2 for SNODAS and SNOTEL by Water Year",
    color = "R2"
  ) +
  theme_classic()
rsquare_snodas_snotel_plot

# plot rmse
rmse_snodas_snotel_plot <- snodas_snotel_stats %>%
  select(waterYear, rmse_depth, rmse_swe) %>% 
  pivot_longer(c(rmse_depth, rmse_swe),
               names_to = "rmse", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rmse)) +
  geom_point(aes(color = rmse)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 1), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Root Mean Square Error",
    title = "RMSE for SNODAS and SNOTEL by Water Year",
    color = "RMSE"
  ) +
  theme_classic()
rmse_snodas_snotel_plot

grid.arrange(rsquare_snodas_snotel_plot, rmse_snodas_snotel_plot, ncol = 1)
```

```{r, echo = FALSE}
# compartive stats for snodas vs snotel
stats_snotel <- snodas_snotel_stats %>% 
  select(waterYear, contains("snotel")) %>% 
  select(-snotel_depth_mm, -snotel_swe_mm) %>% 
  rename(variance_depth = var_snotel_depth,
         variance_swe = var_snotel_swe,
         median_depth = med_snotel_depth,
         median_swe = med_snotel_swe,
         st_dev_depth = sd_snotel_depth,
         st_dev_swe = sd_snotel_swe) %>% 
  mutate(observation = "SNOTEL")

stats_snodas <- snodas_snotel_stats %>% 
  select(waterYear, contains("snodas")) %>% 
  select(-snodas_snow_depth_mm, -snodas_swe_mm) %>% 
  rename(variance_depth = var_snodas_depth,
         variance_swe = var_snodas_swe,
         median_depth = med_snodas_depth,
         median_swe = med_snodas_swe,
         st_dev_depth = sd_snodas_depth,
         st_dev_swe = sd_snodas_swe) %>% 
  mutate(observation = "SNODAS")
  

stats_dist_snodas_snotel <- rbind(stats_snotel, stats_snodas)

### SWE ###
variance_swe_plot <- ggplot(stats_dist_snodas_snotel, aes( x = waterYear, y = variance_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Variance",
    title = "Variance for Daily Winter SWE (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

sd_swe_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = st_dev_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Standard Deviation",
    title = "Standard Deviation of Daily Winter SWE (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

median_swe_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = median_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Median of Daily Winter SWE"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()
  
  
grid.arrange(median_swe_plot, sd_swe_plot, variance_swe_plot, ncol = 1)


### snow depth ###
variance_depth_plot <- ggplot(stats_dist_snodas_snotel, aes( x = waterYear, y = variance_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Variance",
    title = "Variance of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

sd_depth_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = st_dev_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Standard Deviation",
    title = "Standard Deviation of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

median_depth_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = median_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Median of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()
  
  
grid.arrange(median_depth_plot, sd_depth_plot, variance_depth_plot, ncol = 1)
```



Timing of max swe

```{r, echo = FALSE}
max_swe <- wc_snodas_snotel %>% 
  addWaterYear() 
  
# snodas
max_swe_snodas_time <- max_swe %>% 
  select(Date, waterYear, snodas_swe_mm) %>% 
  group_by(waterYear) %>% 
  top_n(1, snodas_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snodas_swe_mm) %>% 
  mutate(type = "snodas")

# SNOTEL
max_swe_snotel_time <- max_swe %>% 
  select(Date, waterYear, snotel_swe_mm) %>% 
  filter(waterYear >= 2010) %>% 
  group_by(waterYear) %>% 
  top_n(1, snotel_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snotel_swe_mm) %>% 
  mutate(type = "snotel")

# check actual dates of SNOTEL max
snotel_max_hist <- max_swe_snotel_time %>% 
  ungroup() %>% 
  mutate(month_day = paste0(month(Date), "-", day(Date))) %>% 
  count(month_day) %>% 
  ggplot(aes(x = month_day)) +
  geom_histogram()


max_swe_all <- rbind(max_swe_snodas_time, max_swe_snotel_time)

# plot
ggplot(max_swe_all, aes(x = waterYear, y= water_week)) +
  geom_point(aes(color = type)) +
  geom_line(aes(color = type)) +
  labs(
    title = "Timing of Peak SWE by Water Week for Whiskey Creek Station",
    x = "Water Year",
    y = "Week of Water Year"
  ) +
  theme_bw()

```



### Monthly max from Nov-Apr

```{r, echo = FALSE}
wc_snodas_snotel_monthly <- wc_snodas_snotel_wint %>% 
  group_by(waterYear, month) %>% 
  summarize(snodas_snow_depth_mm = max(snodas_snow_depth_mm),
            snodas_swe_mm = max(snodas_swe_mm),
            snotel_depth_mm = max(snotel_depth_mm),
            snotel_swe_mm = max(snotel_swe_mm)) 
```

```{r, echo = FALSE}
# correlations
ggplot(wc_snodas_snotel_monthly, aes(x = snodas_swe_mm, y = snotel_swe_mm)) +
  geom_point(aes(color = waterYear)) +
  geom_smooth(method = lm) +
  labs(
    title = "SNOTEL and SNODAS Maximum Monthly SWE Relationship"
  ) +
  theme_classic()


```



```{r, echo = FALSE}
# correlations
ggplot(wc_snodas_snotel_monthly, aes(x = snodas_snow_depth_mm, y = snotel_depth_mm)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(
    title = "SNOTEL and SNODAS Monthly maximum snow depth Relationship"
  ) +
  theme_classic()


```

#### Comparison with PRISM data

```{r, echo = FALSE}
# read in whiskey creek prism  precip data
wc_prism_precip <- read_csv("../Data/wc_prism_precip.csv")

# whiskey creek prism tmean data
wc_prism_tmean <- read_csv("../Data/wc_prism_tmean.csv")

# combine precip adn temp
wc_prism_precip_tmean <- left_join(wc_prism_precip, wc_prism_tmean, by = "Date") %>% 
  addWaterYear()
```

```{r, echo = FALSE}

# set the seasons
winter = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr")
summer = c("May", "Jun", "Jul", "Aug", "Sep")
winter_wet = c("Dec", "Jan", "Feb", "Mar")
spring_dry =  c("Apr", "May", "Jun")
summer_wet = c("Jul", "Aug", "Sep")

# show how much total precipitation this area gets every month over the years
wc_prism_precip_months <- wc_prism_precip %>% 
  mutate(year = year(Date), month = month(Date, label=TRUE,abbr=TRUE)) %>% 
  addWaterYear() %>% 
  group_by(waterYear, month) %>% 
  summarize(
    monthly_precip_mm = sum(ppt)
  ) %>% 
  mutate(bimodal_season = ifelse(month %in% summer, "summer",
                                 "winter")) %>% 
  mutate(wet_dry = ifelse(month %in% winter_wet, "winter_wet",
                          ifelse(month %in% spring_dry, "spring_dry",
                                 ifelse(month %in% summer_wet, "summer_wet",
                                        "fall_dry"))))
  
# plot monthly total precip
ggplot(wc_prism_precip_months, aes(x = waterYear, y = monthly_precip_mm)) +
  geom_col() +
  facet_wrap(~factor(month, wy_order)) +
  labs(
    x = "Water Year",
    y = "Total Precipitation (mm)",
    title = "Whiskey Creek Total PRISM Precipitation By Month"
  ) +
  theme_bw()

# plot total precip separated by winter/summer
wc_prism_precip_months %>% 
  group_by(waterYear, bimodal_season) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
ggplot(aes(x = waterYear, y = season_precip_mm)) +
  geom_col(aes(fill = bimodal_season)) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Whiskey Creek Total PRISM Precipitation By Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

# plot total precip separated by wet/dry seasons
wc_prism_precip_months %>% 
  group_by(waterYear, wet_dry) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
ggplot(aes(x = waterYear, y = season_precip_mm)) +
  geom_col(aes(fill = wet_dry)) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Whiskey Creek Total PRISM Precipitation By Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

# plot total precip separated by wet/dry seasons percent
wc_prism_precip_months %>% 
  group_by(waterYear, wet_dry) %>% 
  summarize(season_precip_mm = sum(monthly_precip_mm)) %>% 
  filter(waterYear != 2020) %>% 
  mutate(percent = season_precip_mm/sum(season_precip_mm)) %>% 
ggplot(aes(x = waterYear, y = percent)) +
  geom_col(aes(fill = wet_dry)) +
  labs(
    x = "Year",
    y = "Fraction of Total Precipitation",
    title = "Whiskey Creek Total PRISM Precipitation By Wet and Dry Season"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1981,2019, by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

```

#### freezing days
```{r, echo = FALSE}
# determine when snow and when rain
wc_prism_freeze <- wc_prism_precip_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  mutate(precip_type = ifelse(tmean <= 0, "snow", "rain")) %>% 
  filter(precip_type != is.na(precip_type)) %>% 
  mutate(prism_snow_mm = ifelse(precip_type == "snow", ppt, 0))

# prism snow accumulation
wc_prism_snow_all <- wc_prism_freeze %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>% 
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(5,6,7,8,9), 0,  prism_snow_mm_accum))

# dialy estimated snofall
ggplot(wc_prism_snow_all, aes(x=Date, y = prism_snow_mm_accum)) +
  geom_line() +
  labs(
    x = "Year",
    y = "PRISM Estimated Snowfall Accumulation (mm)",
    title = "PRISM Estimated Snowfall Accumulation"
  ) +
  scale_x_date(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

# maximum accumulated snow per year
wc_prism_max <- wc_prism_snow_all %>% 
  group_by(waterYear) %>% 
  summarize(prism_snow_mm_accum = max(prism_snow_mm_accum))

# plot
ggplot(wc_prism_max, aes(x = waterYear, y = prism_snow_mm_accum)) +
  geom_col()

# anomaly for 1981-2001 normal
wc_prism_max_anomaly_91_01 <- wc_prism_max %>% 
  filter(waterYear <= 2001) %>% 
  ungroup() %>% 
  summarize(av = mean(prism_snow_mm_accum))

# anomaly for 1981-2010 normal
wc_prism_max_anomaly_91_10 <- wc_prism_max %>% 
  filter(waterYear <= 2010) %>% 
  ungroup() %>% 
  summarize(av = mean(prism_snow_mm_accum))

wc_prism_max_anomaly <- wc_prism_max %>% 
  mutate(av_81_01 = wc_prism_max_anomaly_91_01$av) %>% 
  mutate(av_81_10 = wc_prism_max_anomaly_91_10$av) %>% 
  mutate(anomaly_81_01 = (prism_snow_mm_accum - av_81_01)/prism_snow_mm_accum) %>%
  mutate(anomaly_81_10 = (prism_snow_mm_accum - av_81_10)/prism_snow_mm_accum)

# anomaly for 1981-2001 normal
ggplot(wc_prism_max_anomaly, aes(x = waterYear, y = anomaly_81_01)) +
  geom_col()

# anomaly for 1981-2010 normal
ggplot(wc_prism_max_anomaly, aes(x = waterYear, y = anomaly_81_10)) +
  geom_col()



```

```{r, echo = FALSE}

wc_prism_snowpack <- wc_prism_snow_all %>% 
  mutate(decade = ifelse(waterYear %in% c(1981:1990), "1981-1990", 
                         ifelse(waterYear %in% c(1991:2000), "1991-2000",
                                ifelse(waterYear %in% c(2001:2010), "2001-2010",
                                       "2011-2020")))) %>% 
  filter(month(Date) %in% c(11,12,1,2,3,4))

ggplot(wc_prism_snowpack, aes(x=Date, y = prism_snow_mm_accum)) +
  geom_line()

# distribution
  ggplot(wc_prism_snowpack, aes(x = prism_snow_mm_accum, 
                                y = factor(decade, 
                                                c("2011-2020", "2001-2010", "1991-2000", "1981-1990")))) +
  geom_density_ridges(scale = 7,
                      aes(fill = decade),
                      size = 0.3,
                      color = "NA",
                      alpha = 0.5,
                      show.legend = FALSE) +
  theme_minimal() +
  labs(
    x = "Snowfall (mm)",
    y = "Decade"
  )

```


```{r, echo = FALSE}
# how much total snow vs rain per month
wc_prism_snow <- wc_prism_freeze %>% 
  filter(waterYear != 1981) %>% 
  filter(waterYear != 2020) %>% 
  mutate(month = month(Date)) %>% 
  group_by(waterYear, precip_type) %>% 
  summarize(
    precip_mm = sum(ppt)
  ) %>% 
  mutate(percent = (precip_mm/sum(precip_mm)*100)) 

# how much rain vs snow
ggplot(wc_prism_snow, aes(x = waterYear, y = precip_mm)) +
  geom_point(aes(color = precip_type))+
  geom_line(aes(color = precip_type)) +
  geom_smooth(method = "lm", se = FALSE, aes(color = precip_type)) +
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    color = "Type of Precipitation",
    title = "Total PRISM Precipitation Per Water Year Whiskey Creek "
  ) +
  theme_classic()

ggplot(wc_prism_snow, aes(x = waterYear, y = precip_mm)) +
  geom_col(aes(fill = precip_type))+
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    color = "Type of Precipitation",
    title = "Total PRISM Precipitation Per Water Year Whiskey Creek "
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1982,2019,by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

# percent rain vs snow
ggplot(wc_prism_snow, aes(x = waterYear, y = percent)) +
  geom_col(aes(fill = precip_type)) +
  labs(
    x = "Water Year",
    y = "Percent of Days",
    fill = "Precipitation Type",
    title = "Percent of Precipitation Each Water Year Falling as Rain or Snow"
  ) +
  scale_x_continuous(expand = c(0,0), breaks = seq(1982,2019,by = 5)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()
```

Temperature of hottest average temperature days each water year

```{r, echo = FALSE}
# hottest day each water year
whiskey_hot <- wc_prism_precip_tmean %>% 
  group_by(waterYear) %>% 
  summarize(tmean = max(tmean, na.rm = TRUE)) %>% 
  filter(waterYear != 2020) # not a complete water year

# graphit
graph_season_trend(whiskey_hot, whiskey_hot$tmean, whiskey_hot$waterYear, 
                   title = "Average Temperature of Hottest Day Each Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Temperature C") +
  geom_point()

# hottest day each winter water year winter = nov, dec, jan, feb, mar, apr
whiskey_hot_winter <- wc_prism_precip_tmean %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11, 12, 1, 2, 3, 4)) %>% 
  group_by(waterYear) %>% 
  summarize(tmean = max(tmean, na.rm = TRUE))

# graphit
graph_season_trend(whiskey_hot_winter, whiskey_hot_winter$tmean, whiskey_hot_winter$waterYear, 
                   title = "Average Temperature of Hottest Day Each Winter", 
                   xlabel = "Winter Water Year",
                   ylabel = "Temperature C")  

```

Number of freezing days, when average temperature is at or below 0

```{r, echo = FALSE}
# number of freezing days per water year
# remove 1981 because wasn't a full winter (data starts in Jan)
num_freeze <- wc_prism_precip_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  group_by(waterYear) %>% 
  summarize(ndayfr=sum(freeze, na.rm = TRUE)) %>% 
  filter(waterYear != 1981)

# graphit
graph_season_trend(num_freeze, num_freeze$ndayfr, num_freeze$waterYear, 
                   title = "Number of Freezing Days per Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Number of Freezing Days")  
```

freezing days correlation to SNOTEL swe max
```{r, echo = FALSE}

snotel_swe_max <- wc_clean_metric %>% 
  addWaterYear() %>% 
  filter(waterYear != 2009) %>% 
  group_by(waterYear) %>% 
  summarize(swe_mm = max(swe_mm, na.rm = TRUE)) %>% 
  left_join(num_freeze, by = "waterYear")

ggplot(snotel_swe_max, aes(x = ndayfr, y = swe_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Number of PRISM Freezing Days",
    y = "Max SNOTEL SWE (mm)"
  ) +
  theme_bw() +
  theme(text = element_text(size=14))

# theres a relationship between number of freezing days and maximum SWE
cor.test(snotel_swe_max$swe_mm, snotel_swe_max$ndayfr)
summary(lm(snotel_swe_max$swe_mm ~ snotel_swe_max$ndayfr))

```

Freezing days relationship to week of water year of max swe

```{r, echo = FALSE}
# not much of a relationship
snotel_swe_max_week <- left_join(max_swe_snotel_time, num_freeze, by = "waterYear")

ggplot(snotel_swe_max_week, aes(x = ndayfr, y = water_week)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

```{r, echo = FALSE}
# dataframe of just prism snow accumulation
wc_prism_snow_accum <- wc_prism_freeze %>% 
  select(-waterYear) %>% 
  addWaterYear() %>% 
  mutate(month = month(Date)) %>% 
  drop_na(ppt) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(4,5,6,7,8,9,10), 0, 
                                      prism_snow_mm_accum)) 
```


```{r, echo = FALSE}
# snotel data is accumulated precipitation, so we need to sum prism snowfall to estimate how much snow
# combine with snotel data & make precipitaitona ccumulation columsn for prism data
wc_prism_snotel <- wc_clean_metric %>% 
  select(Date, "snotel_ppt_mm" = precip_mm, 
         "snotel_snow_depth_mm" = snow_depth_mm, "snotel_swe_mm" = swe_mm,
         "snotel_av_temp" = air_temp_av_C) %>% 
  left_join(wc_prism_freeze, by = "Date")  %>% 
  select(-waterYear) %>% 
  addWaterYear() %>% 
  mutate(month = month(Date)) %>% 
  drop_na(ppt) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(4,5,6,7,8,9,10), 0, 
                                      prism_snow_mm_accum))
```

```{r, echo = FALSE}
# set colors for graphing prism, snodas, snotel
prism_snotel_snodas_colors = c("tomato3", "green4", "steelblue3")
```


```{r, echo = FALSE}
# look at snotel, prism, and snodas daily all together
wc_prism_snotel_snodas <- left_join(wc_prism_snotel, wc_snodas_swe, by = "Date") %>% 
  select(Date, waterYear, snotel_av_temp, snotel_ppt_mm, snotel_swe_mm, "prism_precip" = ppt,
         "prism_av_temp" = tmean, prism_snow_mm_accum, snodas_swe_mm)

wc_prism_snotel_snodas_swe_tidy <- wc_prism_snotel_snodas %>% 
  select(Date, waterYear, "SNOTEL SWE (mm)" = snotel_swe_mm, 
         "PRISM Accumulated Snowfall (mm)" = prism_snow_mm_accum, 
        "SNODAS SWE (mm)" = snodas_swe_mm) %>% 
  pivot_longer(c("SNOTEL SWE (mm)", "PRISM Accumulated Snowfall (mm)","SNODAS SWE (mm)"),
               values_to = "snow_mm", 
               names_to = "measurement") %>% 
  filter(Date > "2009-09-01")
  


```


```{r, echo = FALSE}
# daily overlay
ggplot(wc_prism_snotel_snodas_swe_tidy, aes(x = Date, y = snow_mm)) +
  geom_line(aes(color = measurement)) +
  labs(
    x = "Date",
    y = "milimeters",
    color = " "
  ) +
  scale_color_manual(values = prism_snotel_snodas_colors) +
  scale_y_continuous(expand = c(0,0)) +
 # scale_x_date(expand = c(0,0), date_breaks = "2 years", date_minor_breaks = "1 year") +
  scale_x_date(expand = c(0,0), breaks = as.Date(c("2010-01-01", "2012-01-01", "2014-01-01",
                                                   "2016-01-01", "2018-01-01", "2020-01-01"))) +
  theme_classic() +
  theme(legend.position="bottom", text = element_text(size=14)) 

# snotel, prism, and snodas stats
wc_prism_snotel_snodas_stats <- wc_prism_snotel_snodas_swe_tidy %>% 
  ungroup() %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  rename(data = measurement) %>% 
  mutate(data = factor(data, c("SNOTEL SWE (mm)", "SNODAS SWE (mm)", 
                               "PRISM Accumulated Snowfall (mm)"))) %>% 
  group_by(data) %>% 
  summarize(
    median = round(median(snow_mm, na.rm = TRUE),2),
    mean = round(mean(snow_mm, na.rm = TRUE),2),
    variance = round(var(snow_mm, na.rm = TRUE),2),
    st_dev = round(sd(snow_mm, na.rm = TRUE),2)
  )

wc_prism_snotel_snodas_table <- kable(wc_prism_snotel_snodas_stats,
                                      col.names = c("Data", "Median", "Mean", "Variance", "SD")) %>% 
  kable_styling(bootstrap_option = c("striped", "hover"), full_width = F)

wc_prism_snotel_snodas_table
```




```{r, echo = FALSE}

# maximum values for each water year 
wc_prism_snotel_snodas_swe_max <- wc_prism_snotel_snodas_swe_tidy %>% 
  filter(waterYear != 2009) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  group_by(waterYear, measurement) %>% 
  top_n(1, snow_mm) %>% 
  top_n(1, Date)

# plot it 

ggplot(wc_prism_snotel_snodas_swe_max, aes(x = waterYear, y = snow_mm)) +
  geom_line(aes(color = measurement)) +
  geom_point(aes(color = measurement)) +
  labs(
    x = "Water Year",
    y = "milimeters",
    color = " "
  ) +
  scale_color_manual(values = prism_snotel_snodas_colors) +
  scale_x_continuous(breaks = seq(2010,2020, by=2)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(legend.position="bottom", text = element_text(size=14)) 

# Max SWE correlation graph 
wc_prism_snotel_snodas_swe_max_cor  <- wc_prism_snotel_snodas_swe_max %>% 
  select(waterYear, measurement, snow_mm) %>% 
  pivot_wider(names_from = measurement, values_from = snow_mm) %>% 
  pivot_longer(c("SNODAS SWE (mm)", "PRISM Accumulated Snowfall (mm)"), 
               names_to = "Snow Product", values_to = "snow_mm") %>% 
  clean_names()

ggplot(wc_prism_snotel_snodas_swe_max_cor, aes(x = snotel_swe_mm, y = snow_mm)) +
  geom_point(aes(color = snow_product), show.legend = FALSE) +
  geom_smooth(method = "lm", se = FALSE, aes(color = snow_product), show.legend = FALSE) +
  labs(
    x = "Max SNOTEL SWE (mm)",
    y = "milimeters", 
    color = "Remote Sensing Product"
  ) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 

wc_prism_max_lm <- wc_prism_snotel_snodas_swe_max_cor %>% 
  filter(snow_product == "PRISM Accumulated Snowfall (mm)") %>% 
  ungroup() %>% 
  mutate(snodas_estimate = ifelse(snotel_swe_mm >= snow_mm, "underestimate", "overestimate"))

wc_snodas_max_lm <- wc_prism_snotel_snodas_swe_max_cor %>% 
  filter(snow_product == "SNODAS SWE (mm)") %>% 
  ungroup() %>% 
  mutate(snodas_estimate = ifelse(snotel_swe_mm >= snow_mm, "underestimate", "overestimate"))

summary(lm(wc_prism_max_lm$snow_mm ~ wc_prism_max_lm$snotel_swe_mm))
summary(lm(wc_snodas_max_lm$snow_mm ~ wc_snodas_max_lm$snotel_swe_mm))
rmse(wc_snodas_max_lm$snow_mm, wc_snodas_max_lm$snotel_swe_mm)

```


```{r, echo = FALSE}
# relationship between SNOTEL average temp and PRISM average temp

ggplot(wc_prism_snotel, aes(x = snotel_av_temp, y = tmean)) +
  geom_point()

# correlation
cor.test(wc_prism_snotel$snotel_av_temp, wc_prism_snotel$tmean)
```

```{r, echo = FALSE}
# look at correlations for winter months
wc_prism_snotel_wint <- wc_prism_snotel %>% 
  filter(month %in% c(11,12,1,2,3)) 

# graph estimated accumulated prism snow with snotel swe
ggplot(wc_prism_snotel, aes(x = Date, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism")) +
  labs(
    y = "milimeters",
    title = "PRISM Estimated Snow Accumulation and SNOTEL SWE"
  ) +
  scale_x_date(date_breaks = "2 years", expand = c(0,0)) +
  theme_classic()

# root mean square error
rmse(wc_prism_snotel_wint$prism_snow_mm_accum, wc_prism_snotel_wint$snotel_swe_mm)

summary(lm(wc_prism_snotel_wint$prism_snow_mm_accum ~ wc_prism_snotel_wint$snotel_swe_mm))

```

```{r, echo = FALSE}
# root mean square error
rmse(wc_prism_snotel_wint$prism_snow_mm_accum, wc_prism_snotel_wint$snotel_swe_mm)

# RMSE each year

wc_rmse_clean <- wc_prism_snotel_wint %>% 
  ungroup() %>% 
  select(Date, waterYear, snotel_swe_mm, prism_snow_mm_accum)

wc_rmse_wy <- wc_rmse_clean %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)

# RMSE each month
wc_rmse_month <- wc_rmse_clean %>% 
  mutate(month = month(Date)) %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)
```


```{r, echo = FALSE}
# plot it
wc_prism_snotel_wint <- wc_prism_snotel_wint %>% 
  mutate(month_name = month(Date, label = TRUE, abbr=TRUE))

wc_prism_snotel_wint_plot <- graph_correlation(wc_prism_snotel_wint, x_data = 
                                                 wc_prism_snotel_wint$prism_snow_mm_accum, 
                        y_data = wc_prism_snotel_wint$snotel_swe_mm,
                        title = "Whiskey Creek SNOTEL and PRISM", 
                         xlabel = "PRISM Estimated Snow Accumulation (mm)",
                         ylabel = "SNOTEL SWE (mm)") +
                          geom_point(aes(color = month_name)) +
                          scale_color_manual(values = colors)
wc_prism_snotel_wint_plot

# by each month
wc_prism_snotel_wint_plot + 
  facet_wrap(~factor(month_name, level = c("Nov", "Dec", "Jan", "Feb", "Mar"))) +
  labs(color = "month")
  
# by water year
wc_prism_snotel_wint_plot + facet_wrap(~waterYear) +
  labs(color = "month")

# rmse and mae per water year
wc_prism_snotel_wint_rmse_mae <- wc_prism_snotel_wint %>% 
  select(waterYear, snotel_swe_mm, prism_snow_mm_accum) %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  mutate(mae = mae(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)

# rmse and mae per month
wc_prism_snotel_wint_rmse_mae_mnth <- wc_prism_snotel_wint %>% 
  ungroup() %>% 
  select(Date, snotel_swe_mm, prism_snow_mm_accum) %>% 
  mutate(month = month(Date)) %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  mutate(mae = mae(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)
  


```


```{r, echo = FALSE}
# max comparison
wc_prism_snotel_max <- wc_prism_snotel %>% 
  filter(waterYear != 2009) %>% 
  summarize(
    snotel_swe_mm = max(snotel_swe_mm, na.rm = TRUE),
    prism_snow_mm_accum = max(prism_snow_mm_accum)
  ) %>% 
  left_join(num_freeze, by = "waterYear")

# graph estimated accumulated prism snow with snotel swe
ggplot(wc_prism_snotel_max, aes(x = waterYear, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel_swe_mm")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism_snow_mm")) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "milimeters",
    colour = "Type",
    title = "PRISM Estimated Max Accumulated Snowfall and SNOTEL Max SWE"
  ) +
  theme_classic()

# r squared
print("Adjusted R2 value is")
summary(lm(wc_prism_snotel_max$prism_snow_mm_accum ~ wc_prism_snotel_max$snotel_swe_mm))$adj.r.squared
```


#### precipitation comparison 

```{r, echo = FALSE, eval = FALSE}

ggplot(wc_prism_snotel, aes(x = snotel_ppt_mm, y = prism_ppt_mm)) +
  geom_point() +
  labs(
    x = "Snotel precipitation (mm)",
    y = "PRISM precipitation (mm)"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

```


### Beaver Spring

* Station ID:	1143

* State:	Arizona

* Network:	SNOTEL

* County:	Apache

* Elevation:	9200 ft.

* Latitude:	36.33

* Longitude:	-109.06

* HUC:	140802040401

```{r, echo = FALSE}
# grab snotel data from NRCS
beaver_spring_snotel <- grabNRCS.data(network = "SNTL",
                                      site_id = 1143,
                                      timescale = "daily",
                                      DayBgn = "2009-01-01",
                                      DayEnd = "2020-06-01")

# clean dataset
bs_clean <- beaver_spring_snotel %>% 
  select(Date,
          "air_temp_av_F" = Air.Temperature.Average..degF., 
         "air_temp_max_F" = Air.Temperature.Maximum..degF., 
        "air_temp_min_F" =  Air.Temperature.Minimum..degF.,
        "precip_in" = Precipitation.Accumulation..in..Start.of.Day.Values,
        "snow_depth_in" = Snow.Depth..in..Start.of.Day.Values,
        "swe_in" = Snow.Water.Equivalent..in..Start.of.Day.Values) %>% 
  mutate(Date = ymd(Date))

# make everything metric
# convert everything to metric units
bs_clean_metric <- bs_clean %>% 
  mutate(air_temp_av_C = round(conv_unit(air_temp_av_F, "F", "C"),2)) %>% 
  mutate(air_temp_max_C = round(conv_unit(air_temp_max_F, "F", "C"),2)) %>% 
  mutate(air_temp_min_C = round(conv_unit(air_temp_min_F, "F", "C"),2)) %>% 
  mutate(precip_mm = conv_unit(precip_in, "inch", "mm")) %>% 
  mutate(snow_depth_mm = conv_unit(snow_depth_in, "inch", "mm")) %>%
  mutate(swe_mm = conv_unit(swe_in, "inch", "mm")) %>% 
  select(Date, air_temp_av_C, air_temp_max_C, air_temp_min_C, precip_mm, snow_depth_mm, swe_mm) %>% 
  mutate(station = "beaver_springs")
```


### First explore NAs
```{r, echo = FALSE}
# table summary of missing variables
miss_var_summary(bs_clean_metric)

# plot of missing variables
vis_miss(bs_clean_metric, sort = TRUE)

# visualize NA intersections across variables with an upset plot
gg_miss_upset(bs_clean_metric, 
              nsets = 7)
```

##### Daily basic trends

```{r, echo = FALSE}
# SWE
bs_swe_daily <- ggplot(bs_clean, aes(x = Date, y = swe_in)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (in)",
                      title = "Daily SWE"
                    ) +
                    theme_classic()

# snow depth
bs_depth_daily <- ggplot(bs_clean, aes(x = Date, y = snow_depth_in)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (in)",
                      title = "Daily Snow Depth"
                    ) +
                    theme_classic()

# max temp
bs_max_temp_daily <- ggplot(bs_clean, aes(x = Date, y = air_temp_max_F)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (F)",
                      title = "Daily Maximum Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(bs_swe_daily, bs_depth_daily, bs_max_temp_daily, ncol = 2)

```



#### Monthly basic trends

```{r, echo = FALSE}
# maximum monthly data
bs_max_monthly <- bs_clean_metric %>% 
  group_by(year = year(Date), 
           month = month(Date)) %>% 
  summarize(air_temp_av_C = max(air_temp_av_C, na.rm = TRUE),
            air_temp_max_C = max(air_temp_max_C, na.rm = TRUE),
            air_temp_min_C = max(air_temp_min_C, na.rm = TRUE),
            swe_mm = max(swe_mm, na.rm = TRUE),
            snow_depth_mm = max(snow_depth_mm, na.rm = TRUE)) %>% 
  mutate(Date = ymd(paste0(year, "-", month, "-01")))

# SWE
bs_swe_monthly <- ggplot(bs_max_monthly, aes(x = Date, y = swe_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "SWE (mm)",
                      title = "Max Monthly SWE"
                    ) +
                    theme_classic()

# snow depth
bs_depth_monthly <- ggplot(bs_max_monthly, aes(x = Date, y = snow_depth_mm)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Snow Depth (mm)",
                      title = "Max Monthly Snow Depth"
                    ) +
                    theme_classic()

# max temp
bs_max_temp_monthly <- ggplot(bs_max_monthly,
                              aes(x = Date, y = air_temp_max_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Maximum Air Temperature"
                    ) +
                    theme_classic()


# max average temp
bs_max_av_temp_monthly <- ggplot(bs_max_monthly,
                              aes(x = Date, y = air_temp_av_C)) +
                    geom_line() +
                    labs(
                      x = "Date",
                      y = "Air Temperature (C)",
                      title = "Max Monthly Average Air Temperature"
                    ) +
                    theme_classic()

# show all together
grid.arrange(bs_swe_monthly, bs_depth_monthly, bs_max_temp_monthly, bs_max_av_temp_monthly, ncol = 2)


```

## Whiskey creek vs Beaver springs
```{r, echo = FALSE}
# stack the two snotel station data
wc_bs <- wc_clean_metric %>% 
  mutate(station = "whiskey_creek") %>% 
  rbind(bs_clean_metric) %>% 
  addWaterYear() %>% 
  filter(waterYear > 2009)

# SWE
wc_bs_swe_plot <- ggplot(wc_bs, aes(x = Date, y = swe_mm)) +
                geom_line(aes(color = station)) +
                labs(
                  y = "SWE (mm)", 
                  x = "Date",
                  color = "Station"
                ) +
                scale_y_continuous(expand = c(0,0)) +
                scale_x_date(expand = c(0,0), date_breaks = "2 years")+
                theme_bw()

# Snow depth
wc_bs_depth_plot <-ggplot(wc_bs, aes(x = Date, y = snow_depth_mm)) +
                  geom_line(aes(color = station)) +
                  labs(
                    y = "Snow Depth (mm)", 
                    x = "Date",
                    color = "Station"
                  ) +
                  scale_y_continuous(expand = c(0,0)) +
                  scale_x_date(expand = c(0,0), date_breaks = "2 years")+
                  theme_bw()


grid.arrange(wc_bs_swe_plot, wc_bs_depth_plot, ncol = 1)
```

#### Seasonal trends for both snotel sites
```{r, echo = FALSE}
# clean for seasonal calculations
wc_bs_swe_seasonal <- wc_bs %>% 
  select(Date, waterYear, swe_mm, snow_depth_mm, station) %>% 
  mutate(month = month(Date)) %>% 
  mutate(day = day(Date)) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  filter(month %in% c(1,2,3,4,11,12)) 

# daily average swe
wc_bs_swe_seasonal_daily <- wc_bs_swe_seasonal %>% 
            group_by(station, month, day) %>% 
            summarize(
              mean_swe = mean(swe_mm, na.rm = TRUE),
              mean_depth = mean(snow_depth_mm, na.rm = TRUE)
            ) %>% 
            mutate(fake_md = paste0(month,"-", day)) %>% 
            mutate(fake_year = ifelse(month %in% c(11,12), 1999, 2000)) %>% 
            mutate(fake_date = mdy(paste0(fake_md, "-", fake_year)))

# average daily swe from 2009-2020
ggplot(wc_bs_swe_seasonal_daily, aes(x = fake_date, y = mean_swe)) +
  geom_line(aes(color = station)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  labs(
    x = "Month",
    y = "Mean SWE (mm)"
  )

# weekly average SWE
wc_bs_swe_seasonal_week <- wc_bs_swe_seasonal %>% 
  group_by(station, water_week) %>% 
  summarize(
    mean_swe = mean(swe_mm, na.rm = TRUE),
    mean_depth = mean(snow_depth_mm, na.rm = TRUE)
  )

# average weekly swe from 2009-2020
ggplot(wc_bs_swe_seasonal_week, aes(x = water_week, y = mean_swe)) +
  geom_line(aes(color = station)) +
  labs(
    x = "Week of Water Year",
    y = "Mean SWE (mm)"
  ) +
  theme_classic()

# monthly median SWE
wc_bs_swe_seasonal %>% 
  mutate(month_lab = month(Date, label = TRUE, abbr = TRUE)) %>% 
  group_by(station, month_lab) %>% 
  summarize(swe_mm = median(swe_mm, na.rm = TRUE)) %>% 
  ggplot(aes(x = factor(month_lab, wy_order), y = swe_mm, group = station)) +
  geom_point(aes(color = station)) +
  geom_line(aes(color = station)) +
  labs(
    x = "Month",
    y = "SWE (mm)",
    color = "Station"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()
  
```


```{r, echo = FALSE}
# look at the correlation scatter plot between the swe and depth
wc_bs_swe_depth <- wc_clean_metric %>% 
  select(Date, "swe_mm_wc" = swe_mm, "snow_depth_mm_wc" = snow_depth_mm) %>% 
  left_join(bs_clean_metric, by = "Date") %>% 
  select(Date, swe_mm_wc, snow_depth_mm_wc, "swe_mm_bs" = swe_mm, "snow_depth_mm_bs" = snow_depth_mm)

 wc_bs_swe_corr <- ggplot(wc_bs_swe_depth, aes( x = swe_mm_wc, y = swe_mm_bs)) +
          geom_point() +
          geom_smooth(method= "lm", se = FALSE) +
          labs(
            x = "Whiskey Creek SWE (mm)",
            y = "Beaver Springs SWE (mm)"
          ) +
          theme_bw()
 
wc_bs_depth_corr <- ggplot(wc_bs_swe_depth, aes( x = snow_depth_mm_wc, y = snow_depth_mm_bs)) +
          geom_point() +
          geom_smooth(method= "lm", se = FALSE) +
          labs(
            x = "Whiskey Creek Snow Depth (mm)",
            y = "Beaver Springs Snow Depth (mm)"
          ) +
          theme_bw()

grid.arrange(wc_bs_swe_corr, wc_bs_depth_corr, ncol = 1)
```



### timing of peak swe

```{r, echo = FALSE}
ws_bs_peak_swe <- wc_bs %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  ungroup() %>% 
  select(waterYear, water_week, swe_mm, station) %>% 
  group_by(station, waterYear) %>% 
  top_n(1, swe_mm) %>% 
  top_n(1, water_week)

ggplot(ws_bs_peak_swe, aes(x = waterYear, y = water_week)) +
  geom_point(aes(color = station)) +
  geom_line(aes(color = station)) +
  labs(
    x = "Water Year",
    y = "Water Week"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =2)) +
  theme_classic()

```



### SNOTEL vs SNODAS

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# read in snodas swe
bs_snodas_swe <- read_csv("../Data/beaver_springs_snodas_swe.csv",
                          col_names = c("Date", "snodas_swe_mm"), 
                          skip = 1)
# read in snodas snow depth
bs_snodas_depth <- read_csv("../Data/beaver_springs_snodas_depth.csv",
                          col_names = c("Date", "snodas_snow_depth_mm"), 
                          skip = 1)
```

#### Daily data

```{r, echo = FALSE}
# combine snodas and snotel
bs_snodas_snotel <- left_join(bs_snodas_swe, bs_snodas_depth) %>% 
  left_join(wc_clean_metric, by = "Date") %>% 
  select(Date, snodas_swe_mm,  snodas_snow_depth_mm,
         "snotel_depth_mm" = snow_depth_mm, 
         "snotel_swe_mm" = swe_mm) %>% 
  addWaterYear()

# tidy the data for plotting
bs_tidy_snodas_snotel <- bs_snodas_snotel %>% 
  pivot_longer(c(snodas_swe_mm, snodas_snow_depth_mm, 
                 snotel_depth_mm, snotel_swe_mm),
               names_to = "type", values_to = "value_mm") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
### SWE ###
bs_das_tel_swe <- bs_tidy_snodas_snotel %>% 
  filter(type == "snodas_swe_mm" | type ==  "snotel_swe_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "SWE (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL SWE Beaver Springs"
  ) +
  theme_classic()
wc_das_tel_swe

# clean the data for graphing
bs_snodas_snotel_wint <- bs_snodas_snotel %>% 
  mutate(month = month(Date)) %>% 
  mutate(month_name = month(Date, label = TRUE, abbr = TRUE)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  filter(waterYear > 2009)# 2009 not a full water year

# rmse and mae by month between snodas swe and snotel swe
bs_snodas_snotel_wint_rmse_mae <- bs_snodas_snotel_wint %>% 
  ungroup() %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(snodas_swe_mm, snotel_swe_mm)) %>% 
  mutate(mae = mae(snodas_swe_mm, snotel_swe_mm)) %>% 
  select(month_name, rmse, mae) %>% 
  slice(1)

# rmse and mae by water year between snodas swe and snotel swe
bs_snodas_snotel_wint_rmse_mae_wy <- bs_snodas_snotel_wint %>% 
  ungroup() %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(snodas_swe_mm, snotel_swe_mm)) %>% 
  mutate(mae = mae(snodas_swe_mm, snotel_swe_mm)) %>% 
  select(waterYear, rmse, mae) %>% 
  slice(1)
  
# all    
all_bs_snodas <-  graph_correlation(bs_snodas_snotel_wint, x_data = bs_snodas_snotel_wint$snodas_swe_mm, 
                        y_data = bs_snodas_snotel_wint$snotel_swe_mm,
                        title = "Beaver Spring SNOTEL and SNODAS Daily SWE", 
                         xlabel = "SNODAS SWE (mm)",
                         ylabel = "SNOTEL SWE (mm)") +
                          geom_point(aes(color = month_name)) +
                          scale_color_manual(values = colors)
 
all_bs_snodas

# by water year
all_bs_snodas + facet_wrap(~waterYear) +
  labs(color = "month")
  
# by month  
all_bs_snodas + facet_wrap(~month_name) 
```

```{r, echo = FALSE}
# snow depth
bs_das_tel_depth <- bs_tidy_snodas_snotel %>% 
  filter(type == "snodas_snow_depth_mm" | type ==  "snotel_depth_mm") %>% 
  ggplot(aes(x = Date, y = value_mm)) +
  geom_line(aes(color = type)) +
  labs(
    y = "Snow depth (mm)",
    type = "Data Source",
    title = "SNODAS and SNOTEL Snow Depth"
  ) +
  scale_x_date(expand = c(0,0)) +
  theme_classic()
bs_das_tel_depth

# correlation
ggplot(bs_snodas_snotel_wint,
       aes(x = snotel_depth_mm, y = snodas_snow_depth_mm)) +
  geom_point(aes(color = waterYear)) +
  geom_smooth(method = lm) +
  labs(
    x = "SNOTEL Snow Depth (mm)",
    y = "SNODAS Snow Depth (mm)",
    title = "SNOTEL and SNODAS Daily Snow depth Relationship"
  ) +
  theme_classic()

# correlations by water year
  ggplot(bs_snodas_snotel_wint, aes(x = snotel_depth_mm, 
                                    y = snodas_snow_depth_mm)) +
      geom_point() +
      geom_smooth(method = lm) +
      facet_wrap(~waterYear) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0,0)) +
      labs(
        x = "SNOTEL Snow Depth (mm)",
        y = "SNODAS Snow Depth (mm)",
        title = "SNOTEL and SNODAS Daily Snow depth Relationship"
      ) +
     theme_classic()
```


```{r, echo = FALSE}
# r squared for each water year for snodas vs snotel swe and depth
rsquare_snodas_snotel_bs <- bs_snodas_snotel_wint %>% 
  ungroup() %>% 
  select(-Date, -month) %>% 
  group_by(waterYear) %>% 
  mutate(rsquare_depth = 
           summary(lm(snodas_snow_depth_mm ~ snotel_depth_mm))$adj.r.squared)  %>% 
  mutate(rsquare_swe = 
            summary(lm(snodas_swe_mm ~ snotel_swe_mm))$adj.r.squared) %>% 
  select(waterYear, rsquare_depth, rsquare_swe) %>% 
  slice(1)

# plot rsquared
rsquare_snodas_snotel_plot_bs <- rsquare_snodas_snotel_bs %>%
  pivot_longer(c(rsquare_depth, rsquare_swe),
               names_to = "rsquared", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rsquared)) +
  geom_point(aes(color = rsquared)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Adjusted R2",
    title = "Adjusted R2 for SNODAS and SNOTEL by Water Year",
    color = "R2"
  ) +
  theme_classic()
rsquare_snodas_snotel_plot_bs

summary(lm(bs_snodas_snotel_wint$snodas_swe_mm ~ bs_snodas_snotel_wint$snotel_swe_mm))


```

```{r, echo = FALSE}
# Stats for snotel and snodas
snodas_snotel_stats <- bs_snodas_snotel_wint %>% 
  ungroup() %>% 
  select(-Date, -month) %>% 
  group_by(waterYear) %>% 
  mutate(rsquare_depth = 
           summary(lm(snodas_snow_depth_mm ~ snotel_depth_mm))$adj.r.squared)  %>% 
  mutate(rsquare_swe = 
            summary(lm(snodas_swe_mm ~ snotel_swe_mm))$adj.r.squared) %>% 
  mutate(rmse_depth = mltools::rmse(snodas_snow_depth_mm, snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(rmse_swe = mltools::rmse(snodas_swe_mm, snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(var_snotel_depth = var(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(var_snodas_depth = var(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(var_snotel_swe = var(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(var_snodas_swe = var(snodas_swe_mm, na.rm = TRUE)) %>% 
  mutate(med_snotel_depth = median(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(med_snodas_depth = median(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(med_snotel_swe = median(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(med_snodas_swe = median(snodas_swe_mm, na.rm = TRUE)) %>% 
  mutate(sd_snotel_depth = sd(snotel_depth_mm, na.rm = TRUE))  %>% 
  mutate(sd_snodas_depth = sd(snodas_snow_depth_mm, na.rm = TRUE))  %>% 
  mutate(sd_snotel_swe = sd(snotel_swe_mm, na.rm = TRUE)) %>% 
  mutate(sd_snodas_swe = sd(snodas_swe_mm, na.rm = TRUE)) %>% 
#  select(waterYear, rsquare_depth, rsquare_swe, var_depth, var_swe, rmse_depth, rmse_swe) %>% 
  slice(1)

rmse(bs_snodas_snotel_wint$snodas_swe_mm, bs_snodas_snotel_wint$snotel_swe_mm)

# plot rsquared
rsquare_snodas_snotel_plot <- snodas_snotel_stats %>%
  select(waterYear, rsquare_depth, rsquare_swe) %>% 
  pivot_longer(c(rsquare_depth, rsquare_swe),
               names_to = "rsquared", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rsquared)) +
  geom_point(aes(color = rsquared)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Adjusted R2",
    title = "Adjusted R2 for SNODAS and SNOTEL by Water Year",
    color = "R2"
  ) +
  theme_classic()
rsquare_snodas_snotel_plot

# plot rmse
rmse_snodas_snotel_plot <- snodas_snotel_stats %>%
  select(waterYear, rmse_depth, rmse_swe) %>% 
  pivot_longer(c(rmse_depth, rmse_swe),
               names_to = "rmse", values_to = "values") %>% 
  ggplot(aes(x = waterYear, y = values)) +
  geom_line(aes(color = rmse)) +
  geom_point(aes(color = rmse)) +
  scale_x_continuous(breaks = seq(2010,2020, by = 1), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Root Mean Square Error",
    title = "RMSE for SNODAS and SNOTEL by Water Year",
    color = "RMSE"
  ) +
  theme_classic()
rmse_snodas_snotel_plot

grid.arrange(rsquare_snodas_snotel_plot, rmse_snodas_snotel_plot, ncol = 1)
```

```{r, echo = FALSE}
# compartive stats for snodas vs snotel
stats_snotel <- snodas_snotel_stats %>% 
  select(waterYear, contains("snotel")) %>% 
  select(-snotel_depth_mm, -snotel_swe_mm) %>% 
  rename(variance_depth = var_snotel_depth,
         variance_swe = var_snotel_swe,
         median_depth = med_snotel_depth,
         median_swe = med_snotel_swe,
         st_dev_depth = sd_snotel_depth,
         st_dev_swe = sd_snotel_swe) %>% 
  mutate(observation = "SNOTEL")

stats_snodas <- snodas_snotel_stats %>% 
  select(waterYear, contains("snodas")) %>% 
  select(-snodas_snow_depth_mm, -snodas_swe_mm) %>% 
  rename(variance_depth = var_snodas_depth,
         variance_swe = var_snodas_swe,
         median_depth = med_snodas_depth,
         median_swe = med_snodas_swe,
         st_dev_depth = sd_snodas_depth,
         st_dev_swe = sd_snodas_swe) %>% 
  mutate(observation = "SNODAS")
  

stats_dist_snodas_snotel <- rbind(stats_snotel, stats_snodas)

### SWE ###
variance_swe_plot <- ggplot(stats_dist_snodas_snotel, aes( x = waterYear, y = variance_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Variance",
    title = "Variance for Daily Winter SWE (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

sd_swe_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = st_dev_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Standard Deviation",
    title = "Standard Deviation of Daily Winter SWE (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

median_swe_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = median_swe)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Median of Daily Winter SWE"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()
  
  
grid.arrange(median_swe_plot, sd_swe_plot, variance_swe_plot, ncol = 1)


### snow depth ###
variance_depth_plot <- ggplot(stats_dist_snodas_snotel, aes( x = waterYear, y = variance_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Variance",
    title = "Variance of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

sd_depth_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = st_dev_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "Standard Deviation",
    title = "Standard Deviation of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()

median_depth_plot <- ggplot(stats_dist_snodas_snotel, aes(x = waterYear, y = median_depth)) +
  geom_line(aes(color = observation)) + 
  geom_point(aes(color = observation)) + 
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Median of Daily Winter Snow Depth (mm)"
  ) +
  scale_x_continuous(breaks = seq(2010,2020, by =1)) +
  theme_bw()
  
  
grid.arrange(median_depth_plot, sd_depth_plot, variance_depth_plot, ncol = 1)
```

```{r, echo = FALSE}
max_swe <- bs_snodas_snotel %>% 
  addWaterYear() 
  
# snodas
max_swe_snodas_time <- max_swe %>% 
  select(Date, waterYear, snodas_swe_mm) %>% 
  group_by(waterYear) %>% 
  top_n(1, snodas_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snodas_swe_mm) %>% 
  mutate(type = "snodas")

# SNOTEL
max_swe_snotel_time <- max_swe %>% 
  select(Date, waterYear, snotel_swe_mm) %>% 
  filter(waterYear >= 2010) %>% 
  group_by(waterYear) %>% 
  top_n(1, snotel_swe_mm) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  top_n(1, water_week) %>% 
  rename(swe_mm = snotel_swe_mm) %>% 
  mutate(type = "snotel")

max_swe_all <- rbind(max_swe_snodas_time, max_swe_snotel_time)

# plot
ggplot(max_swe_all, aes(x = waterYear, y= water_week)) +
  geom_point(aes(color = type)) +
  geom_line(aes(color = type)) +
  labs(
    title = "Timing of Peak SWE by Water Week for Beaver Springs Station",
    x = "Water Year",
    y = "Week of Water Year"
  ) +
  theme_bw()

```

### Comparison with PRISM data

```{r, echo = FALSE}
# read in beaver springs prism precipiation data
bs_prism_precip <- read_csv("../Data/bs_prism_precip.csv")

# read in beaver springs prism temperature data
bs_prism_tmean <- read_csv("../Data/bs_prism_tmean.csv")

# combine precip adn temp
bs_prism_precip_tmean <- left_join(bs_prism_precip, bs_prism_tmean, by = "Date") %>% 
  addWaterYear()
```

```{r, echo = FALSE}
# show how much total precipitation this area gets every month over the years
bs_prism_precip_months <- bs_prism_precip %>% 
  mutate(year = year(Date), month = month(Date, label = TRUE, abbr = TRUE)) %>% 
  group_by(year, month) %>% 
  summarize(
    monthly_precip_mm = sum(ppt)
  )

ggplot(bs_prism_precip_months, aes(x = year, y = monthly_precip_mm)) +
  geom_col() +
  geom_smooth(method = "lm", se = FALSE)+
  facet_wrap(~month) +
  labs(
    x = "Year",
    y = "Total Precipitation (mm)",
    title = "Beaver Springs Total PRISM Precipitation By Month"
  ) +
  facet_wrap(~factor(month, c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep"))) +
  theme_bw()
```

#### freezing days
```{r, echo = FALSE}

bs_prism_freeze <- bs_prism_precip_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  mutate(precip_type = ifelse(tmean <= 0, "snow", "rain")) %>% 
  filter(precip_type != is.na(precip_type)) %>% 
  mutate(prism_snow_mm = ifelse(precip_type == "snow", ppt, 0))

# how much total snow vs rain per month
bs_prism_snow <- bs_prism_freeze %>% 
  filter(waterYear != 1981) %>% 
  mutate(month = month(Date)) %>% 
  group_by(waterYear, precip_type) %>% 
  summarize(
    precip_mm = sum(ppt)
  )%>% 
  mutate(percent = (precip_mm/sum(precip_mm)*100))

# how much rain vs snow
ggplot(bs_prism_snow, aes(x = waterYear, y = precip_mm)) +
  geom_point(aes(color = precip_type))+
  geom_line(aes(color = precip_type)) +
  geom_smooth(method = "lm", se = FALSE, aes(color = precip_type)) +
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    color = "Type of Precipitation",
    title = "Total PRISM Precipitation Per Water Year Beaver Springs"
  ) +
  theme_classic()

# percentrain vs snow
ggplot(bs_prism_snow, aes(x = waterYear, y = percent)) +
  geom_col(aes(fill = precip_type))+
  labs(
    x = "Water Year",
    y = "Precipitation (mm)",
    color = "Type of Precipitation",
    title = "Percent Precipitation Per Water Year Beaver Springs"
  ) +
  theme_classic()
```




Temperature of hottest average temperature days each water year

```{r, echo = FALSE}
# hottest day each water year
beaver_hot <- bs_prism_precip_tmean %>% 
  group_by(waterYear) %>% 
  summarize(tmean = max(tmean, na.rm = TRUE)) %>% 
  filter(waterYear != 2020) # not a complete water year

# graphit
graph_season_trend(beaver_hot, beaver_hot$tmean, beaver_hot$waterYear, 
                   title = "Average Temperature of Hottest Day Each Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Temperature C")  

# hottest day each winter water year winter = nov, dec, jan, feb, mar, apr
beaver_hot_winter <- bs_prism_precip_tmean %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11, 12, 1, 2, 3, 4)) %>% 
  group_by(waterYear) %>% 
  summarize(tmean = max(tmean, na.rm = TRUE))

# graphit
graph_season_trend(beaver_hot_winter, beaver_hot_winter$tmean, beaver_hot_winter$waterYear, 
                   title = "Average Temperature of Hottest Day Each Winter", 
                   xlabel = "Winter Water Year",
                   ylabel = "Temperature C")  

```

Number of freezing days, when average temperature is at or below 0

```{r, echo = FALSE}
# number of freezing days per water year
# remove 1981 because wasn't a full winter (data starts in Jan)
num_freeze <- bs_prism_precip_tmean %>% 
  mutate(freeze = ifelse(tmean <= 0, 1, 0)) %>% 
  group_by(waterYear) %>% 
  summarize(ndayfr=sum(freeze, na.rm = TRUE)) %>% 
  filter(waterYear != 1981)

# graphit
graph_season_trend(num_freeze, num_freeze$ndayfr, num_freeze$waterYear, 
                   title = "Number of Freezing Days per Water Year", 
                   xlabel = "Water Year",
                   ylabel = "Number of Freezing Days")  
```

```{r, ehco = FALSE}
# raw prism snowfall
bs_prism_snow <- bs_prism_freeze %>% 
  select(-waterYear) %>% 
  addWaterYear() %>% 
  mutate(month = month(Date)) %>% 
  drop_na(ppt) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(4,5,6,7,8,9,10), 0, 
                                      prism_snow_mm_accum))
```


freezing days correlation to SNOTEL swe max
```{r, echo = FALSE}

snotel_swe_max <- bs_clean_metric %>% 
  addWaterYear() %>% 
  filter(waterYear != 2009) %>% 
  group_by(waterYear) %>% 
  summarize(swe_mm = max(swe_mm, na.rm = TRUE)) %>% 
  left_join(num_freeze, by = "waterYear")

ggplot(snotel_swe_max, aes(x = ndayfr, y = swe_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Number of PRISM Freezing Days",
    y = "SNOTEL SWE (mm)"
  ) +
  theme_bw() +
  theme(text = element_text(size=14))

# theres a relationship between number of freezing days and maximum SWE
cor.test(snotel_swe_max$swe_mm, snotel_swe_max$ndayfr)
summary(lm(snotel_swe_max$swe_mm ~ snotel_swe_max$ndayfr))

```

#### PRISM Snow accumulation vs SNOTEL SWE

```{r, echo = FALSE}
# snotel data is accumulated precipitation, so we need to sum prism snowfall to estimate how much snow
# combine with snotel data & make precipitaitona ccumulation columsn for prism data
bs_prism_snotel <- bs_clean_metric %>% 
  select(Date, "snotel_ppt_mm" = precip_mm, 
         "snotel_snow_depth_mm" = snow_depth_mm, "snotel_swe_mm" = swe_mm,
         "snotel_av_temp" = air_temp_av_C) %>% 
  left_join(bs_prism_freeze, by = "Date")  %>% 
  select(-waterYear) %>% 
  addWaterYear() %>% 
  mutate(month = month(Date)) %>% 
  drop_na(ppt) %>% 
  group_by(waterYear) %>% 
  mutate(prism_snow_mm_accum = cumsum(prism_snow_mm)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum = ifelse(month(Date) %in% c(4,5,6,7,8,9,10), 0, 
                                      prism_snow_mm_accum))

# look at correlations for winter months
bs_prism_snotel_wint <- bs_prism_snotel %>% 
  filter(month %in% c(11,12,1,2,3)) 

# graph estimated accumulated prism snow with snotel swe
ggplot(bs_prism_snotel, aes(x = Date, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism")) +
  labs(
    y = "milimeters",
    title = "PRISM Estimated Snow Accumulation and SNOTEL SWE"
  ) +
  scale_x_date(date_breaks = "2 years", expand = c(0,0)) +
  theme_classic()

summary(lm(bs_prism_snotel_wint$prism_snow_mm_accum ~ bs_prism_snotel_wint$snotel_swe_mm))
```

```{r, echo = FALSE}
# look at snotel, prism, and snodas daily all together
bs_prism_snotel_snodas <- left_join(bs_prism_snotel, bs_snodas_swe, by = "Date") %>% 
  select(Date, waterYear, snotel_av_temp, snotel_ppt_mm, snotel_swe_mm, "prism_precip" = ppt,
         "prism_av_temp" = tmean, prism_snow_mm_accum, snodas_swe_mm)

bs_prism_snotel_snodas_swe_tidy <- bs_prism_snotel_snodas %>% 
  select(Date, waterYear, "SNOTEL SWE (mm)" = snotel_swe_mm, 
         "PRISM Accumulated Snowfall (mm)" = prism_snow_mm_accum, 
        "SNODAS SWE (mm)" = snodas_swe_mm) %>% 
  pivot_longer(c("SNOTEL SWE (mm)", "PRISM Accumulated Snowfall (mm)","SNODAS SWE (mm)"),
               values_to = "snow_mm", 
               names_to = "measurement") %>% 
  filter(Date > "2009-09-01")
  
# snotel, prism, and snodas stats
bs_prism_snotel_snodas_stats <- bs_prism_snotel_snodas_swe_tidy %>% 
  ungroup() %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11,12,1,2,3,4)) %>% 
  rename(data = measurement) %>% 
  mutate(data = factor(data, c("SNOTEL SWE (mm)", "SNODAS SWE (mm)", 
                               "PRISM Accumulated Snowfall (mm)"))) %>% 
  group_by(data) %>% 
  summarize(
    median = round(median(snow_mm, na.rm = TRUE),2),
    mean = round(mean(snow_mm, na.rm = TRUE),2),
    variance = round(var(snow_mm, na.rm = TRUE),2),
    st_dev = round(sd(snow_mm, na.rm = TRUE),2)
  )

bs_prism_snotel_snodas_table <- kable(bs_prism_snotel_snodas_stats,
                                      col.names = c("Data", "Median", "Mean", "Variance", "SD")) %>% 
  kable_styling(bootstrap_option = c("striped", "hover"), full_width = F)

bs_prism_snotel_snodas_table
```


```{r, echo = FALSE}
# daily overlay
ggplot(bs_prism_snotel_snodas_swe_tidy, aes(x = Date, y = snow_mm)) +
  geom_line(aes(color = measurement)) +
  labs(
    x = "Date",
    y = "milimeters",
    color = " "
  ) +
  scale_color_manual(values = prism_snotel_snodas_colors) +
  scale_y_continuous(expand = c(0,0)) +
 # scale_x_date(expand = c(0,0), date_breaks = "2 years", date_minor_breaks = "1 year") +
  scale_x_date(expand = c(0,0), breaks = as.Date(c("2010-01-01", "2012-01-01", "2014-01-01",
                                                   "2016-01-01", "2018-01-01", "2020-01-01"))) +
  theme_classic() +
  theme(legend.position="bottom", text = element_text(size=14)) 

```


```{r, echo = FALSE}
# maximum values for each water year 
bs_prism_snotel_snodas_swe_max <- bs_prism_snotel_snodas_swe_tidy %>% 
  filter(waterYear != 2009) %>% 
  mutate(water_week = to_water_week(Date)) %>% 
  group_by(waterYear, measurement) %>% 
  top_n(1, snow_mm) %>% 
  top_n(1, Date)

# plot it 

ggplot(bs_prism_snotel_snodas_swe_max, aes(x = waterYear, y = snow_mm)) +
  geom_line(aes(color = measurement)) +
  geom_point(aes(color = measurement)) +
  labs(
    x = "Water Year",
    y = "milimeters",
    color = " "
  ) +
  scale_color_manual(values = prism_snotel_snodas_colors) +
  scale_x_continuous(breaks = seq(2010,2020, by=2)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(legend.position="bottom", text = element_text(size=14)) 

# Max SWE correlation graph 
bs_prism_snotel_snodas_swe_max_cor  <- bs_prism_snotel_snodas_swe_max %>% 
  select(waterYear, measurement, snow_mm) %>% 
  pivot_wider(names_from = measurement, values_from = snow_mm) %>% 
  pivot_longer(c("SNODAS SWE (mm)", "PRISM Accumulated Snowfall (mm)"), 
               names_to = "Snow Product", values_to = "snow_mm") %>% 
  clean_names()

ggplot(bs_prism_snotel_snodas_swe_max_cor, aes(x = snotel_swe_mm, y = snow_mm)) +
  geom_point(aes(color = snow_product), show.legend = FALSE) +
  geom_smooth(method = "lm", se = FALSE, aes(color = snow_product), show.legend = FALSE) +
  labs(
    x = "Max SNOTEL SWE (mm)",
    y = "milimeters", 
    color = "Remote Sensing Product"
  ) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 

bs_prism_max_lm <- bs_prism_snotel_snodas_swe_max_cor %>% 
  filter(snow_product == "PRISM Accumulated Snowfall (mm)")

bs_snodas_max_lm <- bs_prism_snotel_snodas_swe_max_cor %>% 
  filter(snow_product == "SNODAS SWE (mm)") %>% 
  ungroup() %>% 
  mutate(snodas_estimate = ifelse(snotel_swe_mm > snow_mm, "under", "over"))

rmse(bs_prism_max_lm$snow_mm, bs_prism_max_lm$snotel_swe_mm)

summary(lm(bs_prism_max_lm$snow_mm ~ bs_prism_max_lm$snotel_swe_mm))
summary(lm(bs_snodas_max_lm$snow_mm ~ bs_snodas_max_lm$snotel_swe_mm))
  
```


```{r, echo = FALSE}

# plot it
bs_prism_snotel_wint <- bs_prism_snotel_wint %>% 
  mutate(month_name = month(Date, label = TRUE, abbr=TRUE))

bs_prism_snotel_wint_plot <- graph_correlation(bs_prism_snotel_wint, x_data = 
                                                 bs_prism_snotel_wint$prism_snow_mm_accum, 
                        y_data = bs_prism_snotel_wint$snotel_swe_mm,
                        title = "Beaver Spring SNOTEL and PRISM", 
                         xlabel = "PRISM Estimated Snow Accumulation (mm)",
                         ylabel = "SNOTEL SWE (mm)") +
                          geom_point(aes(color = month_name)) +
                          scale_color_manual(values = colors)
bs_prism_snotel_wint_plot

# by each month
bs_prism_snotel_wint_plot + 
  facet_wrap(~factor(month_name, level = c("Nov", "Dec", "Jan", "Feb", "Mar"))) +
  labs(color = "month")
  
# by water year
bs_prism_snotel_wint_plot + facet_wrap(~waterYear) +
  labs(color = "month")

# rmse and mae per water year
bs_prism_snotel_wint_rmse_mae <- bs_prism_snotel_wint %>% 
  select(waterYear, snotel_swe_mm, prism_snow_mm_accum) %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  mutate(mae = mae(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)

# rmse and mae per month
bs_prism_snotel_wint_rmse_mae_mnth <- bs_prism_snotel_wint %>% 
  ungroup() %>% 
  select(Date, snotel_swe_mm, prism_snow_mm_accum) %>% 
  mutate(month = month(Date)) %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  mutate(mae = mae(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)
  


```

```{r, echo = FALSE}
# root mean square error
rmse(bs_prism_snotel_wint$prism_snow_mm_accum, bs_prism_snotel_wint$snotel_swe_mm)

# RMSE each year

bs_rmse_clean <- bs_prism_snotel_wint %>% 
  ungroup() %>% 
  select(Date, waterYear, snotel_swe_mm, prism_snow_mm_accum)

bs_rmse_wy <- bs_rmse_clean %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)

# RMSE each month
bs_rmse_month <- bs_rmse_clean %>% 
  mutate(month = month(Date)) %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(snotel_swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)
```


```{r, echo = FALSE}
# max comparison
bs_prism_snotel_max <- bs_prism_snotel %>% 
  filter(waterYear != 2009) %>% 
  summarize(
    snotel_swe_mm = max(snotel_swe_mm, na.rm = TRUE),
    prism_snow_mm_accum = max(prism_snow_mm_accum)
  )

# graph estimated accumulated prism snow with snotel swe
ggplot(bs_prism_snotel_max, aes(x = waterYear, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel_swe_mm")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism_snow_mm")) +
  scale_x_continuous(breaks = seq(2010,2020, by = 2), expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "milimeters",
    colour = "Type",
    title = "PRISM Estimated Max Accumulated Snowfall and SNOTEL Max SWE"
  ) +
  theme_classic()

# r squared
print("Adjusted R2 value is")
summary(lm(bs_prism_snotel_max$prism_snow_mm_accum ~ bs_prism_snotel_max$snotel_swe_mm))$adj.r.squared
```

#### PRISM snow accumulation using SNOTEL temperature data

```{r, echo = FALSE}

bs_prism_snotel_temp <- bs_prism_snotel %>% 
  mutate(precip_type_snotel_tmean = ifelse(snotel_av_temp <= 0, "snow", "rain")) %>% 
  filter(precip_type_snotel_tmean != is.na(precip_type_snotel_tmean)) %>%
  mutate(prism_snow_mm_snotel = ifelse(precip_type_snotel_tmean == "snow", ppt, 0)) %>% 
  mutate(prism_snow_mm_accum_snotel = cumsum(prism_snow_mm_snotel)) %>%  # accumulate prism precip
  mutate(prism_snow_mm_accum_snotel = ifelse(month(Date) %in% c(5,6,7,8,9), 0,  prism_snow_mm_accum_snotel))

# check to see how well they match
plot(bs_prism_snotel_temp$prism_snow_mm, bs_prism_snotel_temp$prism_snow_mm_snotel)

# graph estimated accumulated prism snow with snotel swe
ggplot(bs_prism_snotel_temp, aes(x = Date, y = snotel_swe_mm)) +
  geom_line(aes(color = "snotel")) +
  geom_line(aes(y = prism_snow_mm_accum, color = "prism")) +
  geom_line(aes(y = prism_snow_mm_accum_snotel, color = "prism_snotel_temp"))+
  labs(
    y = "milimeters",
    title = "PRISM Estimated Snow Accumulation and SNOTEL SWE"
  ) +
  scale_x_date(date_breaks = "2 years", expand = c(0,0)) +
  theme_classic()

```


#### precipitation comparison 

```{r, eval = FALSE}

# combine with snotel data
bs_prism_snotel <- bs_prism_precip %>% 
  mutate(Date = ymd(Date)) %>% 
  rename(prism_ppt_mm = ppt) %>% 
  left_join(bs_clean_metric, by = "Date") %>% 
  select(Date, prism_ppt_mm, "snotel_ppt_mm" = precip_mm, snow_depth_mm, swe_mm)


```