---
title: "Snow Course 85-17"
author: "AnnaClaire Marley"
date: "8/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
library(lubridate)
library(janitor)
library(gdata)
library(naniar)
library(measurements)
library(tidyverse)
library(dataRetrieval)
library(Metrics)
library(mltools)
library(gridExtra)
```

```{r, echo = FALSE}
# path to source graphing functions
path_graph <- "../Functions/Graphing/"
path_clean <- "../Functions/Cleaning/"
```

```{r, echo = FALSE}
# source functions
source(paste0(path_graph, "graph_season_trend.R"))
source(paste0(path_graph, "graph_correlation.R"))
source(paste0(path_clean, "calculate_prism_snowfall.R"))

```


```{r, echo = FALSE}
# set water year order
wint_wy_order <- c("Dec", "Jan", "Feb", "Mar", "Apr")

# colors for months
colors = c('#F6BD60', '#7FB069', '#548687', "plum4", "coral2", "slategrey")
```

```{r, echo = FALSE, message=FALSE, warning = FALSE}
# read in snow course data
snow_course_raw <- read_csv("../Data/nn_snow_course.csv") %>% 
  clean_names()

# read in snotel data
wc_snotel <- read_csv("../Data/wc_snotel.csv") %>% 
  clean_names()
bs_snotel <- read_csv("../Data/bs_snotel.csv")

# read in snodas data
wc_snodas <- read_csv("../Data/whiskey_creek_snodas_swe.csv", col_names = c("date", "swe_mm"), skip = 1)
bs_snodas <- read_csv("../Data/beaver_springs_snodas_swe.csv", col_names = c("date", "swe_mm"), skip = 1)
fr_snodas <- read_csv("../Data/fr_snodas.csv", col_names = c("date", "swe_mm"), skip = 1)
bc_snodas <-read_csv("../Data/bowl_canyon_snodas.csv", col_names = c("date", "swe_mm"), skip = 1) # bowl canyon

# read in prism data
wc_prism <- read_csv("../Data/wc_prism_snow_accum.csv")
bs_prism <- read_csv("../Data/bs_prism_snow.csv")
fr_prism <- read_csv("../Data/fr_prism.csv") %>%  # fluted rock 
  mutate(date = mdy(date))


# for combining files
folder <- "/Users/annaclairemarley/Documents/CAARE_2020/Extra_Data/"
fill_pattern <- "fr_temp*"
variable <- "temp"

df = NULL
  for (file in list.files(folder, pattern = fill_pattern)) {
    readFile = read_csv(paste(folder, file, sep="/"), col_names=c("date", variable), skip = 1)
    if (is.null(df)) { # don't merge the first one
      df = readFile
    } else {
      df = rbind(df, readFile) # bind the files together
    }
  }

```

# SNOTEL Seasonality
```{r, echo = FALSE }
# clean beaver
beav <- bs_snotel %>% 
  select(Date, swe_mm) %>% 
  mutate(station_name = "Beaver Springs")

# snotel seasonal graph
snotel_season <- wc_snotel %>% 
  select("Date"=date, swe_mm) %>% 
  mutate(station_name = "Whiskey Creek") %>% 
  rbind(beav) %>% 
  mutate(month = month(Date, abbr = TRUE, label = TRUE)) %>% 
  filter(month %in% c("Nov", "Dec", "Jan", "Feb", "Mar", "Apr")) %>% 
  group_by(month)

ggplot(snotel_season, aes(x = factor(month, 
                        levels = c("Nov","Dec","Jan", "Feb", "Mar", "Apr")), y = swe_mm)) +
  geom_boxplot(aes(fill = month), alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~station_name) +
  labs( 
    x = "Month",
    y = "SWE (mm)"
    ) +
  theme_bw()

# maximum day beaver springs
beav_max <- beav %>% 
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  top_n(1, swe_mm)

# beaver spring and PRISM RMSE and MAE
bs_prism_snotel_error <- left_join(beav, bs_prism, by = "Date") %>% 
  select(Date, waterYear, swe_mm, prism_snow_mm_accum) %>% 
  mutate(month = month(Date)) %>% 
  filter(month %in% c(11,12,1,2,3)) %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(swe_mm, prism_snow_mm_accum, na.rm = TRUE))

```

# Whiskey creek full record
```{r, echo = FALSE}

wc_prism_snotel_snodas <- wc_prism %>% 
  select("date"= Date, waterYear, prism_snow_mm_accum) %>% 
  left_join(wc_snodas, by = "date") %>% 
  rename(snodas_swe_mm = swe_mm) %>% 
  left_join(wc_snotel, by = "date") %>% 
  select(date, waterYear, prism_snow_mm_accum, snodas_swe_mm, "snotel_swe_mm" = swe_mm) %>% 
  pivot_longer(c(prism_snow_mm_accum, snodas_swe_mm, snotel_swe_mm), values_to = "mm", names_to = "model")

ggplot(wc_prism_snotel_snodas, aes(x = date, y = mm)) +
  geom_line(aes(color = model)) +
  labs(
    x = "Year",
    y = "milimeters",
    title = "Whiskey Creek PRISM, SNOTEL, SNODAS",
    color = ""
  ) +
  scale_x_date(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(legend.position="bottom", text = element_text(size=14)) 
```



```{r, echo = FALSE}
snow_course <- snow_course_raw %>% 
  mutate(Date = dmy(date_collected)) %>% 
  mutate(total_depth_mm = conv_unit(total_snow_depth_in, "inch", "mm")) %>% 
  mutate(av_depth_mm = conv_unit(ave_snow_depth_in, "inch", "mm")) %>% 
  mutate(swe_mm = conv_unit(ave_water_content_in, "inch", "mm")) %>% 
  dplyr::select(Date, station_name, swe_mm, av_depth_mm, total_depth_mm) %>% 
  addWaterYear() %>% 
  rename(date = Date)
```

Look at raw data

```{r, echo = FALSE}
ggplot(snow_course, aes(x = date, y = swe_mm)) +
  #geom_point(aes(color = station_name))+
  geom_line(aes(color = station_name))+
  #geom_col(aes(fill = station_name)) +
  facet_wrap(~station_name)
```

- take out White Clay, Wind Rock Valley, Wheatfields, Tsaile II, Tsaile IA, sloping meadows because incomplete records

```{r,echo = FALSE}

snow_course_trim <- snow_course %>% 
  filter(!station_name %in% c("Sloping Meadows", "Tsaile IA", "Tsaile II", "Wheatfields", 
                              "White Clay", "Wind Rock Valley"))

ggplot(snow_course_trim, aes(x = date, y = swe_mm)) +
  #geom_point(aes(color = station_name))+
  geom_line(aes(color = station_name))+
  #geom_col(aes(fill = station_name)) +
  facet_wrap(~station_name)


```

How much data is missing?

```{r, echo = FALSE, warning = FALSE}

missing_WY <- snow_course_trim %>% 
  dplyr::select(waterYear, station_name, swe_mm) %>% 
  group_by(station_name, waterYear) %>% 
  miss_var_summary()

missing_month <- snow_course_trim %>% 
  mutate(month = month(date, abbr = TRUE, label = TRUE)) %>% 
  filter(!is.na(date)) %>% 
  dplyr::select(date, month, station_name, swe_mm) %>% 
  group_by(station_name, month) %>% 
  miss_var_summary()

# plot missing percent WY
ggplot(missing_WY, aes(x = waterYear, y = pct_miss, group = station_name)) +
  geom_col(aes(fill = station_name)) +
  facet_wrap(~station_name) +
  labs(
    x = "Water Year",
    y = "Percent Data Missing",
    title = "Percent data missing each year"
  ) +
  scale_x_continuous(breaks = seq(1975,2020, by = 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 

# plot missing percent WY
ggplot(missing_month, aes(x = month, y = pct_miss, group = station_name)) +
  geom_col(aes(fill = station_name)) +
  facet_wrap(~station_name) +
  labs(
    x = "Water Year",
    y = "Percent Data Missing",
    title = "Percent data missing each month"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 

```

** A lot of data missing in April let's throw that out

```{r, echo = FALSE}

snow_course_trim %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  filter(month != "Apr") %>% 
  group_by(month) %>% 
  ggplot(aes(x = factor(month, 
                        levels = c("Dec","Jan", "Feb", "Mar")), y = swe_mm)) +
  geom_boxplot(aes(color = month), show.legend = FALSE) +
  facet_wrap(~station_name) +
  labs( 
    x = "Month",
    y = "SWE (mm)"
    ) +
  theme_classic()

```


Look at seasonal monthly median

```{r, echo = FALSE}
# by month
seasonal_median <- snow_course_trim %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  group_by(station_name, month) %>% 
  summarize(swe_mm = median(swe_mm, na.rm = TRUE)) %>% 
  drop_na() %>% 
  filter(month != "Apr")


ggplot(seasonal_median, aes(x = factor(month, wint_wy_order), y = swe_mm, group = station_name)) +
  geom_point(aes(color = station_name)) +
  geom_line(aes(color = station_name)) +
  labs(
    x = "Month",
    y = "SWE (mm)",
    color = "Station"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

# by date

date_median <- snow_course_trim %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  mutate(day = day(date)) %>% 
  mutate(month_day = paste0(month, "-", day)) %>% 
  group_by(station_name, month_day) %>% 
  summarize(swe_mm = median(swe_mm, na.rm = TRUE)) %>% 
  drop_na()


ggplot(date_median, aes(x = month_day, y = swe_mm, group = station_name)) +
  geom_point(aes(color = station_name)) +
  geom_line(aes(color = station_name)) +
  labs(
    x = "Date",
    y = "SWE (mm)",
    color = "Station"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()

```

Seasonal monthly mean

```{r, echo = FALSE}
seasonal_mean <- snow_course_trim %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  group_by(station_name, month) %>% 
  summarize(swe_mm = mean(swe_mm, na.rm = TRUE)) %>% 
  drop_na()

ggplot(seasonal_mean, aes(x = factor(month, wint_wy_order), y = swe_mm, group = station_name)) +
  geom_point(aes(color = station_name)) +
  geom_line(aes(color = station_name)) +
  labs(
    x = "Month",
    y = "SWE (mm)",
    title = "Seasonal Mean",
    color = "Station"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic()
```


# look at max each water year

```{r, echo = FALSE}

sc_wy_max <- snow_course_trim %>% 
  group_by(station_name, waterYear) %>% 
  summarize(swe_mm = max(swe_mm, na.rm = TRUE)) 

ggplot(sc_wy_max, aes(x = waterYear, y = swe_mm)) +
  geom_col(aes(fill = station_name), show.legend = FALSE) +
  facet_wrap(~station_name) +
  labs(
    x = "Water Year",
    y = "SWE (mm)",
    title = "Max SWE"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

```

# Beaver Spring and WHiskey Creek seasonality
```{r, echo = FALSE}
bs_snowcourse <- snow_course_trim %>% 
  filter(station_name == "Beaver Spring")

wc_snowcourse <- snow_course_trim %>% 
  filter(station_name == "Whiskey Creek")

wc_bs_sc <- snow_course_trim %>% 
  filter(station_name %in% c("Whiskey Creek", "Beaver Spring")) %>% 
  drop_na()

# plot raw time series BS vs WC
ggplot(wc_bs_sc, aes(x = date, y = swe_mm)) +
  geom_line(aes(color = station_name))

# plot seasonality
wc_bs_sc %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  filter(month != "Apr") %>% 
  group_by(month) %>% 
  ggplot(aes(x = factor(month, 
                        levels = c("Dec","Jan", "Feb", "Mar")), y = swe_mm)) +
  geom_boxplot(aes(fill = month), alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~station_name) +
  labs( 
    x = "Month",
    y = "SWE (mm)"
    ) +
  theme_bw()

# plot maximums
max_sc <- wc_bs_sc %>% 
  group_by(station_name, waterYear) %>% 
  top_n(1, swe_mm) 

ggplot(max_sc, aes(x = waterYear, y = swe_mm)) +
  geom_col(aes(fill = station_name), position = "dodge") +
  scale_x_continuous(expand = c(0,0), breaks= seq(1984,2017, by = 2)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = colors)

# rnak water eyars
max_sc_rank <- max_sc %>% 
  slice(1) %>% 
  ungroup() %>% 
  dplyr::select(waterYear, station_name, swe_mm) %>% 
  pivot_wider(names_from = "station_name", values_from = "swe_mm")

```



# Beaver Spring

## PRISM, snow course, snodas, snotel

```{r, echo = FALSE}
# swe and prism snowfall
bs_sc_snotel_prism_snodas <- bs_prism %>% 
  dplyr::select("date" = Date, prism_snow_mm_accum) %>% 
  left_join(bs_snowcourse, by = "date") %>% 
  dplyr::select(date, "snowcourse" = swe_mm, prism_snow_mm_accum) %>% 
  left_join(bs_snodas, by = "date") %>% 
  rename(snodas = swe_mm) %>% 
  rename(Date = date) %>% 
  left_join(bs_snotel, by = "Date") %>% 
  dplyr::select("date" = Date, snowcourse, "SNOTEL" = swe_mm, "SNODAS" = snodas,
                "PRISM"=prism_snow_mm_accum)

bs_sc_snotel_prism_snodas_plot <- bs_sc_snotel_prism_snodas %>% 
  pivot_longer(c(SNOTEL, SNODAS, PRISM), 
               names_to = "ground_model", values_to = "mm") %>% 
  mutate(Date = date) %>% 
  addWaterYear()

# graph all of them
bs_sc_snotel_prism_snodas_plot %>% 
  filter(waterYear > 2003) %>% 
ggplot( aes(x = date, y = snowcourse)) +
  geom_col(color = "grey", position = "dodge", width = 2) +
  geom_line(aes(y = mm, color = ground_model), size = 1) +
  labs(
    x = "Year",
    y = "milimeters",
    color = ""
  ) +
  scale_x_date(expand = c(0,0), date_breaks = "2 years") +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(text = element_text(size=14), legend.position="bottom")


# max rnaking for snotel, prism, snodas
max_bs_prism_snodas_snotel <- bs_sc_snotel_prism_snodas_plot %>% 
  dplyr::select(-snowcourse, -Date, -date) %>% 
  filter(waterYear > 2009) %>% 
  ungroup() %>% 
  group_by(ground_model, waterYear) %>% 
  mutate(max = max(mm, na.rm = TRUE)) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(ground_model) %>% 
  arrange(ground_model, max) %>% 
  mutate(rank = 1:n()) %>% 
  dplyr::select(-mm, -max) %>% 
  pivot_wider(c(waterYear, ground_model, rank), names_from = "ground_model", values_from = "rank") %>% 
  arrange(waterYear)

```


## plot PRISM with Snow course SWE
```{r, echo = FALSE}


bs_prism_sc <- bs_prism %>% 
  select("date" = Date, prism_snow_mm_accum) %>% 
  mutate(month = month(date)) %>% 
  left_join(bs_snowcourse, by = "date") %>% 
  select(date, waterYear, prism_snow_mm_accum, swe_mm)%>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) 
 # pivot_longer(c(prism_snow_mm_accum, swe_mm), values_to = "snow_mm", names_to = "monitor") 

ggplot(bs_prism_sc, aes(x = date, y = swe_mm)) +
  geom_col(color = "grey68") +
  geom_line(aes(y = prism_snow_mm_accum), color = "#1219cc") +
  labs(
    x = "Year",
    y = "Snow Course SWE (mm)"
  ) +
  scale_y_continuous(sec.axis = sec_axis(~., "PRISM accumulated snowfall (mm)"), expand = c(0,0)) +
  theme_classic() +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
                axis.text.y.right = element_text(color = "#1219cc"),
                axis.title.y.right = element_text(color = "#1219cc"),
        text = element_text(size=14))


```

```{r, echo = FALSE}

bs_prism_sc_wint <- bs_prism_sc %>% 
  filter(month %in% c("Dec", "Jan", "Feb", "Mar")) %>% 
  drop_na() %>% 
  mutate(median_sc = median(swe_mm)) %>% 
  mutate(median_prism = median(prism_snow_mm_accum)) 

bs_prism_sc_rmse_mae <- bs_prism_sc_wint %>% 
  ungroup() %>% 
  group_by(month) %>% 
  mutate(rmse = rmse(swe_mm, prism_snow_mm_accum)) %>% 
  mutate(mae = mae(swe_mm, prism_snow_mm_accum)) %>% 
  slice(1)

graph_correlation(bs_prism_sc_wint, x_data = bs_prism_sc_wint$prism_snow_mm_accum, 
                        y_data = bs_prism_sc_wint$swe_mm,
                        title = "Beaver Spring Snow Course Daily SWE and PRISM ", 
                         xlabel = "PRISM Accumulated Snowfall (mm)",
                         ylabel = "Snow Course SWE (mm)") +
                        scale_y_continuous(expand = c(0,0), limits = c(0,500)) +
                        scale_x_continuous(expand = c(0,0), limits = c(0,500))
```

# plot SNOTEL SWE and PRISM and SNODAS
```{r, echo}
# SNOTEL and PRSIM
bs_snotel_prism <- bs_sc_snotel_prism_snodas %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  filter(month %in% c("Nov", "Dec", "Jan", "Feb", "Mar")) %>% 
  select(date, month, SNOTEL, PRISM) %>% 
  drop_na() %>% 
  mutate(Date = date) %>% 
  addWaterYear()

graph_correlation(bs_snotel_prism, x_data = bs_snotel_prism$PRISM, 
                        y_data = bs_snotel_prism$SNOTEL,
                        title = "Beaver Spring SNOTEL Daily SWE and PRISM ", 
                         xlabel = "PRISM Accumulated Snowfall (mm)",
                         ylabel = "SNOTEL SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,400))+
  scale_x_continuous(expand = c(0,0), limits = c(0,400))+
  facet_wrap(~waterYear)

# SNOTEL and SNODAS
bs_snotel_snodas <- bs_sc_snotel_prism_snodas %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  filter(month %in% c("Nov", "Dec", "Jan", "Feb", "Mar")) %>% 
  select(date, month, SNOTEL, SNODAS) %>% 
  drop_na() %>% 
  mutate(Date = date) %>% 
  addWaterYear()

graph_correlation(bs_snotel_snodas, x_data = bs_snotel_snodas$SNODAS, 
                        y_data = bs_snotel_snodas$SNOTEL,
                        title = "Beaver Spring SNOTEL Daily SWE and SNODAS ", 
                         xlabel = "SNODAS SWE (mm)",
                         ylabel = "SNOTEL SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,500))+
  scale_x_continuous(expand = c(0,0), limits = c(0,500))+
  facet_wrap(~waterYear)
```


# compare max SWE snow course to PRISM max accumulated
```{r, echo = FALSE, error=TRUE}
wc_prism_max_sc <- wc_prism %>% 
  select(-waterYear) %>% # have to do this for some reason
  addWaterYear() %>% 
  group_by(waterYear) %>% 
  summarize(prism_snow_mm_accum = max(prism_snow_mm_accum, na.rm = TRUE)) 

wc_max_prism_sc <- sc_wy_max %>% 
  filter(station_name == "Whiskey Creek") %>% 
  left_join(wc_prism_max_sc, by = "waterYear")

graph_correlation(wc_max_prism_sc, x_data = wc_max_prism_sc$swe_mm, 
                  y_data = wc_max_prism_sc$prism_snow_mm_accum,
                  title = "Whiskey Creek Max Snow Course SWE and PRISM ", 
                   xlabel = "Snow Course SWE (mm)",
                   ylabel = "PRISM Accumulated Snowfall (mm)") 
```
# Whiskey Creek
## plot PRISM with Snow course SWE
```{r, echo = FALSE}


wc_prism_sc <- wc_prism %>% 
  select("date" = Date, prism_snow_mm_accum) %>% 
  mutate(month = month(date)) %>% 
  left_join(wc_snowcourse, by = "date") %>% 
  select(date, waterYear, prism_snow_mm_accum, swe_mm)%>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) 
 # pivot_longer(c(prism_snow_mm_accum, swe_mm), values_to = "snow_mm", names_to = "monitor") 

ggplot(wc_prism_sc, aes(x = date, y = swe_mm)) +
  geom_col(color = "grey68") +
  geom_line(aes(y = prism_snow_mm_accum), color = "#1219cc") +
  labs(
    x = "Year",
    y = "Snow Course SWE (mm)"
  ) +
  scale_y_continuous(sec.axis = sec_axis(~., "PRISM accumulated snowfall (mm)"), expand = c(0,0)) +
  theme_classic() +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
                axis.text.y.right = element_text(color = "#1219cc"),
                axis.title.y.right = element_text(color = "#1219cc"),
        text = element_text(size=14))


```

```{r, echo = FALSE}
# swe and prism snowfall
wc_sc_snotel_prism_snodas <- wc_prism %>% 
  dplyr::select("date" = Date, prism_snow_mm_accum) %>% 
  left_join(wc_snowcourse, by = "date") %>% 
  dplyr::select(date, "snowcourse" = swe_mm, prism_snow_mm_accum) %>% 
  left_join(wc_snodas, by = "date") %>% 
  rename(snodas = swe_mm) %>% 
  left_join(wc_snotel, by = "date") %>% 
  dplyr::select(date, snowcourse, "SNOTEL" = swe_mm, "SNODAS" = snodas,
                "PRISM"=prism_snow_mm_accum)

wc_sc_snotel_prism_snodas_plot <- wc_sc_snotel_prism_snodas %>% 
  pivot_longer(c(SNOTEL, SNODAS, PRISM), 
               names_to = "ground_model", values_to = "mm") %>% 
  mutate(Date = date) %>% 
  addWaterYear()

# graph all of them
wc_sc_snotel_prism_snodas_plot %>% 
  filter(waterYear > 2003) %>% 
ggplot( aes(x = date, y = snowcourse)) +
  geom_col(color = "grey", position = "dodge", width = 2) +
  geom_line(aes(y = mm, color = ground_model), size = 1) +
  labs(
    x = "Year",
    y = "milimeters",
    color = ""
  ) +
  scale_x_date(expand = c(0,0), date_breaks = "2 years") +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(text = element_text(size=14), legend.position="bottom")


# max rnaking for snotel, prism, snodas
max_wc_prism_snodas_snotel <- wc_sc_snotel_prism_snodas_plot %>% 
  dplyr::select(-snowcourse, -Date, -date) %>% 
  filter(waterYear > 2009) %>% 
  ungroup() %>% 
  group_by(ground_model, waterYear) %>% 
  mutate(max = max(mm, na.rm = TRUE)) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(ground_model) %>% 
  arrange(ground_model, max) %>% 
  mutate(rank = 1:n()) %>% 
  dplyr::select(-mm, -max) %>% 
  pivot_wider(c(waterYear, ground_model, rank), names_from = "ground_model", values_from = "rank") %>% 
  arrange(waterYear)

```

```{r, echo = FALSE}
wc_prism_sc_wint <- wc_prism_sc %>% 
  filter(month %in% c("Dec", "Jan", "Feb", "Mar")) %>% 
  drop_na() %>% 
  mutate(median_sc = median(swe_mm)) %>% 
  mutate(median_prism = median(prism_snow_mm_accum))

graph_correlation(wc_prism_sc_wint, x_data = wc_prism_sc_wint$prism_snow_mm_accum, 
                  y_data = wc_prism_sc_wint$swe_mm,
                  title = "Whiskey Creek Snow Course Daily SWE and PRISM ", 
                   xlabel = "PRISM Accumulated Snowfall (mm)",
                   ylabel = "Snow Course SWE (mm)") +
  scale_x_continuous(expand = c(0,0), limits = c(0,500)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,500))
                  


graph_correlation(wc_prism_sc_wint, x_data = wc_prism_sc_wint$swe_mm, 
                  y_data = wc_prism_sc_wint$prism_snow_mm_accum,
                  title = "Whiskey Creek Snow Course Daily SWE and PRISM ", 
                   xlabel = "Snow Course SWE (mm)",
                   ylabel = "PRISM Accumulated Snowfall (mm)") +
                  geom_point(aes(color = month)) +
                  scale_color_manual(values = colors) +
  facet_wrap(~month)

```


# compare SWE snotel to swe snow course for whiskey creek and beaver springs

```{r, echo = FALSE}

# whiskey creek snotel and snow course
wc_snotel_course <- snow_course_trim %>% 
  filter(station_name == "Whiskey Creek") %>% 
  rename(course_swe_mm = swe_mm) %>% 
  left_join(wc_snotel, by = "date") %>% 
  rename(snotel_swe_mm = swe_mm) %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  drop_na()

graph_correlation(wc_snotel_course, x_data = wc_snotel_course$course_swe_mm, 
                  y_data = wc_snotel_course$snotel_swe_mm,
                  title = "Whiskey Creek Snotel and Snow Course Daily SWE", 
                   xlabel = "Snow Course SWE (mm)",
                   ylabel = "SNOTEL SWE (mm)") +
  scale_x_continuous(expand = c(0,0), limits = c(0,500))

graph_correlation(wc_snotel_course, x_data = wc_snotel_course$av_depth_mm, 
                  y_data = wc_snotel_course$snow_depth_mm,
                  title = "Whiskey Creek Snotel and Snow Course Daily Snow Depth", 
                   xlabel = "Snow Course Snow Depth (mm)",
                   ylabel = "SNOTEL Snow Depth (mm)") +
  scale_x_continuous(expand = c(0,0), limits = c(0,1500)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1500))

# Beaver Springs snotel and snow course
bs_snotel_course <- snow_course_trim %>% 
  filter(station_name == "Beaver Spring") %>% 
  rename(Date = date) %>% 
  rename(course_swe_mm = swe_mm) %>% 
  left_join(bs_snotel, by = "Date") %>% 
  rename(snotel_swe_mm = swe_mm) %>% 
  mutate(month = month(Date, label = TRUE, abbr = TRUE)) %>% 
  filter(!is.na(month)) %>% 
  drop_na()

# SWE
graph_correlation(bs_snotel_course, x_data = bs_snotel_course$course_swe_mm, 
                  y_data = bs_snotel_course$snotel_swe_mm,
                  title = "Beaver Springs Snotel and Snow Course Daily SWE", 
                   xlabel = "Snow Course SWE (mm)",
                   ylabel = "SNOTEL SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,500)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,500))

# depth
graph_correlation(bs_snotel_course, x_data = bs_snotel_course$av_depth_mm, 
                  y_data = bs_snotel_course$snow_depth_mm,
                  title = "Beaver Springs Snotel and Snow Course Daily Snow Depth", 
                   xlabel = "Snow Course Snow Depth (mm)",
                   ylabel = "SNOTEL Snow Depth (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,1500))


```

# compare SWE SNODAS to swe snow course for whiskey creek and beaver springs

```{r, echo = FALSE}
# whiskey creek SNODAS and snow course
wc_snodas_course <- snow_course_trim %>% 
  filter(station_name == "Whiskey Creek") %>% 
  rename(course_swe_mm = swe_mm) %>% 
  left_join(wc_snodas, by = "date")  %>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  drop_na()

graph_correlation(wc_snodas_course, x_data = wc_snodas_course$swe_mm, 
                  y_data = wc_snodas_course$course_swe_mm,
                  title = "Whiskey Creek SNODAS and Snow Course Daily SWE", 
                   xlabel = "SNODAS SWE (mm)",
                   ylabel = "Snow Course SWE (mm)") +
                  scale_y_continuous(expand = c(0,0), limits = c(0,500)) +
                  scale_x_continuous(expand = c(0,0), limits = c(0,500)) 


# error each year
wc_snodas_course_error <- wc_snodas_course %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(course_swe_mm, swe_mm)) %>% 
  mutate(mae = mae(course_swe_mm, swe_mm)) %>% 
  slice(1)


# Beaver Springs SNODAS and snow course
bs_snodas_course <- snow_course_trim %>% 
  filter(station_name == "Beaver Spring") %>% 
  rename(course_swe_mm = swe_mm) %>% 
  left_join(bs_snodas, by = "date")%>% 
  mutate(month = month(date, label = TRUE, abbr = TRUE)) %>% 
  filter(!is.na(month)) %>% 
  drop_na()

# SWE
graph_correlation(bs_snodas_course, x_data = bs_snodas_course$swe_mm, 
                  y_data = bs_snodas_course$course_swe_mm,
                  title = "Beaver Spring SNODAS and Snow Course Daily SWE", 
                   xlabel = "SNODAS SWE (mm)",
                   ylabel = "Snow Course SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,500)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,500))

# error each year
bs_snodas_course_error <- bs_snodas_course %>% 
  group_by(waterYear) %>% 
  mutate(rmse = rmse(course_swe_mm, swe_mm)) %>% 
  mutate(mae = mae(course_swe_mm, swe_mm)) %>% 
  slice(1)


```


## Fluted Rock

```{r,echo = FALSE}
# prism and snow course SWE

fr_snowcourse <- snow_course_trim %>% 
  filter(station_name == "Fluted Rock")

fr_prism_snow <- fr_prism %>% 
  mutate(Date = date) %>% 
  addWaterYear() %>% 
  mutate(prism_snow = calculate_prism_snowfall(prism_df = ., temp = temperature, precip = precip_mm))

fr_sc_prism <- fr_snowcourse %>% 
  select(-waterYear) %>% 
  left_join(fr_prism_snow, by = "date") %>% 
  drop_na() %>% 
  mutate(month = month(date, abbr = TRUE, label = TRUE)) %>% 
  filter(month != "Apr") 
  
graph_correlation(fr_sc_prism, x_data = fr_sc_prism$prism_snow, 
                  y_data = fr_sc_prism$swe_mm,
                  title = "Fluted Rock PRISM and Snow Course Daily SWE", 
                   xlabel = "PRISM Accumulated Snowfall (mm)",
                   ylabel = "Snow Course SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,250)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,250))

fr_sc_prism_nm <- fr_sc_prism %>% 
  filter(month != "Mar")
  
graph_correlation(fr_sc_prism_nm, x_data = fr_sc_prism_nm$prism_snow, 
                  y_data = fr_sc_prism_nm$swe_mm,
                  title = "Fluted Rock PRISM and Snow Course Daily SWE without March", 
                   xlabel = "PRISM Accumulated Snowfall (mm)",
                   ylabel = "Snow Course SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,250)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,250))


```

```{r, echo = FALSE}
# snodas and snow course SWE
fr_sc_snodas <- fr_snowcourse %>% 
  rename(sc_swe_mm = swe_mm) %>% 
  left_join(fr_snodas, by = "date") %>% 
  drop_na() %>% 
  mutate(month = month(date, abbr = TRUE, label = TRUE))

graph_correlation(fr_sc_snodas, x_data = fr_sc_snodas$swe_mm, 
                  y_data = fr_sc_snodas$sc_swe_mm,
                  title = "Fluted Rock SNODAS and Snow Course Daily SWE", 
                   xlabel = "SNODAS SWE (mm)",
                   ylabel = "Snow Course SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,250)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,250))


# plot line graph of snowcourse vs snodas
fr_sc_snodas_all <- fr_snodas %>% 
  rename(snodas_swe_mm = swe_mm) %>% 
  left_join(fr_snowcourse, by = "date") %>% 
  rename(sc_swe_mm = swe_mm)


ggplot(fr_sc_snodas_all, aes(x = date, y = sc_swe_mm)) +
  geom_col(color = "grey", position = "dodge", width = 2) +
  geom_line(aes(y = snodas_swe_mm), size = 1) +
  labs(
    x = "Year",
    y = "milimeters",
    color = ""
  ) +
  scale_x_date(expand = c(0,0), date_breaks = "2 years") +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(text = element_text(size=14), legend.position="bottom")

```

## Bowl Canyon

```{r, echo = FALSE}
bc_snowcourse <- snow_course_trim %>% 
  filter(station_name == "Bowl Canyon")

# snodas and snow course SWE
bc_sc_snodas <- bc_snowcourse %>% 
  rename(sc_swe_mm = swe_mm) %>% 
  left_join(bc_snodas, by = "date") %>% 
  drop_na() %>% 
  mutate(month = month(date, abbr = TRUE, label = TRUE))

graph_correlation(bc_sc_snodas, x_data = bc_sc_snodas$swe_mm, 
                  y_data = bc_sc_snodas$sc_swe_mm,
                  title = "Bowl Canyon SNODAS and Snow Course Daily SWE", 
                   xlabel = "SNODAS SWE (mm)",
                   ylabel = "Snow Course SWE (mm)") +
  scale_y_continuous(expand = c(0,0), limits = c(0,250)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,250))


# plot line graph of snowcourse vs snodas
bc_sc_snodas_all <- bc_snodas %>% 
  rename(snodas_swe_mm = swe_mm) %>% 
  left_join(bc_snowcourse, by = "date") %>% 
  rename(sc_swe_mm = swe_mm)


ggplot(bc_sc_snodas_all, aes(x = date, y = sc_swe_mm)) +
  geom_col(color = "grey", position = "dodge", width = 2) +
  geom_line(aes(y = snodas_swe_mm), size = 1) +
  labs(
    x = "Year",
    y = "milimeters",
    color = ""
  ) +
  scale_x_date(expand = c(0,0), date_breaks = "2 years") +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(text = element_text(size=14), legend.position="bottom")

```

