---
title: "Streamflow"
author: "AnnaClaire Marley"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown analyzes streamflow gage station data in the Chuska mountains

```{r, include = FALSE}
# load packages
library(tidyverse)
library(readxl)
library(gdata)
library(lubridate)
library(dataRetrieval)
library(measurements)
library(naniar)
library(gridExtra)
library(corrplot)
library(janitor)
```

```{r, include = FALSE, message = FALSE}
# read in tidy streamflow data
asaayi <- read_csv("../Data/asaayi_stream_tidy.csv")
wc_stream <- read_csv("../Data/wc_stream_tidy.csv")
capt_tom <- read_csv("../Data/capt_tom_tidy.csv")
wheatfields <- read_csv("../Data/wheatfields_tidy.csv")
tsaile <- read_csv("../Data/tsaile_tidy.csv")
lukachai <- read_csv("../Data/lukachai_tidy.csv")
```

```{r, echo = FALSE, warning = FALSE}
# combine all to one dataframe 
all_streams <- rbind(asaayi, wc_stream, capt_tom, wheatfields, tsaile, lukachai) 

# take out estimated water years
all_streams_ne <- all_streams %>% 
  filter(WY != str_detect(WY, letters))

# make a wide format version for correlation plotting
# this version isn't working -- has a bunch of NULLs I can't get rid of
all_streams_ne_wide_null <- all_streams_ne %>% 
  select(date, station, discharge_cmps) %>% 
  pivot_wider(names_from = station, values_from = discharge_cmps) 

# try a different way 
all_streams_ne_wide <- asaayi %>% 
  select(date, discharge_cmps) %>% 
  rename(asaayi = discharge_cmps) %>% 
  left_join(wc_stream, by = "date") %>% 
  select(date, asaayi, "whiskey_creek" = discharge_cmps) %>% 
  left_join(capt_tom, by = "date") %>% 
  select(date, asaayi, whiskey_creek, "captain_tom" = discharge_cmps) %>% 
  left_join(wheatfields, by = "date") %>% 
  select(date, asaayi, whiskey_creek, captain_tom, "wheatfields" = discharge_cmps) %>% 
  left_join(tsaile, by = "date") %>% 
  select(date, asaayi, whiskey_creek, captain_tom, wheatfields, "tsaile" = discharge_cmps) %>% 
  left_join(lukachai, by = "date") %>% 
  select(date, asaayi, whiskey_creek, captain_tom, wheatfields, tsaile, "lukachai" = discharge_cmps)




```

```{r, include = FALSE}
# read in whiskey creek snotel data
wc_snotel <- read_csv("../Data/wc_snotel.csv") %>% 
  clean_names() %>% 
  mutate(swe_m = conv_unit(swe_mm, "mm", "m"))

bs_snotel <- read_csv("../Data/bs_snotel.csv") %>% 
  clean_names() %>% 
  mutate(swe_m = conv_unit(swe_mm, "mm", "m"))

# order of months
wy_order <- c("NOV", "DEC", "JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")
```


## Investigate missing values first

```{r, echo = FALSE, warning = FALSE}
missing_WY <- all_streams_ne %>% 
  select(WY, station, discharge_cmps) %>% 
  group_by(station, WY) %>% 
  miss_var_summary()

missing_month <- all_streams_ne %>% 
  select(month, station, discharge_cmps) %>% 
  group_by(station, month) %>% 
  miss_var_summary()

# plot missing percent WY
ggplot(missing_WY, aes(x = WY, y = pct_miss, group = station)) +
  geom_col(aes(fill = station)) +
  facet_wrap(~station) +
  labs(
    x = "Water Year",
    y = "Percent Data Missing",
    title = "Percent data missing each year"
  ) +
  scale_x_continuous(breaks = seq(1975,2020, by = 5), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 

# plot missing percent each month
ggplot(missing_month, aes(x = factor(month, wy_order), y = pct_miss)) +
  geom_col(aes(fill = station)) +
  facet_wrap(~station) +
  labs(
    x = "Water Year",
    y = "Percent Data Missing",
    title = "Percent data missing Each Month"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()+
  theme(legend.position="bottom") 
```


# Daily time series all stations

```{r, echo = FALSE}
ggplot(all_streams_ne, aes(x = date, y = discharge_cmps)) +
  geom_line(aes(color = station)) +
  labs(
    x = "Year",
    y = bquote('Daily Discharge' ~(m^3/s))
  ) +
  scale_x_date(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 

ggplot(all_streams_ne, aes(x = date, y = discharge_cmps)) +
  geom_line(aes(color = station)) +
  facet_wrap(~station) +
  labs(
    x = "Year",
    y = bquote('Daily Discharge' ~(m^3/s))
  ) +
  scale_x_date(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw() +
  theme(legend.position="bottom", text = element_text(size=14)) 
```

#### Observations

- Lukachai has the higest discharge, but isn't directly in the Chuskas
- Captain Tom Wash has the next highest discharge


# Monthly peak streamflow

```{r, echo = FALSE}

peak_discharge_month <- all_streams_ne %>% 
  group_by(station, WY, month) %>% 
  summarize(discharge_cmps = max(discharge_cmps, na.rm = TRUE)) %>% 
  filter(discharge_cmps != "-Inf")

# function to graph peak monthly discharge each water year
peak_month_discharge = function(df, station_name = "") {
  
  plot <- df %>% 
    filter(station == station_name) %>% 
    ggplot(aes(x = WY, y = discharge_cmps)) +
    geom_col() +
    facet_wrap(~factor(month, levels=wy_order)) +
    labs(
       x = "Water Year",
       y = bquote('Daily Discharge' ~(m^3/s)),
       title = sprintf("%s Station Peak Monthly Discharge", station_name)
    ) +
    scale_x_continuous(breaks = seq(1980,2020,by = 5)) +
    scale_y_continuous(expand = c(0,0)) +
    theme_bw()
  
  return(plot)
}

peak_month_discharge(peak_discharge_month, station_name = "Asaayi")
peak_month_discharge(peak_discharge_month, station_name = "Captain Tom Wash")
peak_month_discharge(peak_discharge_month, station_name = "Lukachai Creek")
peak_month_discharge(peak_discharge_month, station_name = "Tsaile Creek")
peak_month_discharge(peak_discharge_month, station_name = "Wheatfields")
peak_month_discharge(peak_discharge_month, station_name = "Whiskey Creek")

```


All of the station usually have peak streamflow in April EXCEPT Lukachai, which has it's peak streamflow in August. Maybe Lukachai depends more on summer monsoons while the others are more dominated by snowpack.



# Seasonal streamflow

```{r, echo = FALSE}

# mean
mean_seasonal <- all_streams_ne %>% 
  group_by(station, month) %>% 
  summarize(discharge_cmps = mean(discharge_cmps, na.rm = TRUE))

mean_seasonal_plot <- ggplot(mean_seasonal, aes(x = factor(month, wy_order),
                                                  y = discharge_cmps, group = station)) +
                          geom_line(aes(color = station)) +
                          labs(
                            x = "Month",
                            y = bquote('Discharge' ~(m^3/s)),
                            title = "Mean Monthly Discharge"
                          ) +
                          scale_y_continuous(expand = c(0,0)) +
                          theme_classic() +
                          theme(legend.position="bottom", text = element_text(size=14)) 
# median
median_seasonal <- all_streams_ne %>% 
  group_by(station, month) %>% 
  summarize(discharge_cmps = median(discharge_cmps, na.rm = TRUE))


median_seasonal_plot <- ggplot(median_seasonal, aes(x = factor(month, wy_order),
                                                  y = discharge_cmps, group = station)) +
                          geom_line(aes(color = station)) +
                          labs(
                            x = "Month",
                            y = bquote('Discharge' ~(m^3/s)),
                            title = "Median Monthly Discharge"
                          ) +
                          scale_y_continuous(expand = c(0,0)) +
                          theme_classic() +
                          theme(legend.position="bottom", text = element_text(size=14)) 

grid.arrange(mean_seasonal_plot, median_seasonal_plot, ncol= 1)

```

#### Observations

Mean seasonal monthly discharge shows the expected increase in flow in the summer for Lukachai and Captain Tom was having the higest discharge in the spring. However, this changes when we look at median discharge. This may mean Lukachai's flow is more dominated by Monsoon activity than snowpack.

# Correlations between stations

```{r, echo = FALSE}

# *****Dropped a lot of NAs so be careful about this******
cor_stream <- all_streams_ne_wide %>% 
  select(-date) %>% 
  drop_na() %>% 
  cor()

corrplot(cor_stream, method = "square", type = "upper", 
         addCoef.col = "white", tl.col = "black", tl.srt=45)


```


## compare to Whiskey Creek Snotel Station

###  compare time series of snotel swe and streamflow
```{r, echo = FALSE}
# plot the time series togethr

plot_daily_snotel_steam = function(snotel_df, snotel_name = "", streamflow_df, stream_name = "") {
  
  plot <- streamflow_df %>% 
      filter(station == stream_name) %>% 
      left_join(snotel_df, by = "date") %>% 
      filter(WY > 2009) %>% 
      ggplot(aes(x = date, y = swe_m)) +
        geom_line(size = 1) +
        geom_line(aes(y = discharge_cmps), color = "#1219cc") +
        scale_x_date(expand = c(0,0), limits = ) +
        labs(
          y = "SWE (m)",
          title = sprintf("%s Steamflow Gague and %s SNOTEL SWE", stream_name,  snotel_name)
        ) +
        scale_y_continuous(
          sec.axis = sec_axis(~., "Streamflow Discharge (m3/s)"),
          expand = c(0,0)
        )+
        theme_bw() +
         theme(axis.line.y.right = element_line(color = "#1219cc"),
                axis.text.y.right = element_text(color = "#1219cc"),
                axis.title.y.right = element_text(color = "#1219cc")) 
  
  return(plot)
  
}

# Whiskey Creek Snotel
plot_daily_snotel_steam(snotel_df = wc_snotel, snotel_name = "Whiskey Creek", 
                        streamflow_df = all_streams_ne, stream_name = "Asaayi")
plot_daily_snotel_steam(snotel_df = wc_snotel, snotel_name = "Whiskey Creek", 
                        streamflow_df = all_streams_ne, stream_name = "Whiskey Creek")
plot_daily_snotel_steam(snotel_df = wc_snotel, snotel_name = "Whiskey Creek", 
                        streamflow_df = all_streams_ne, stream_name = "Captain Tom Wash")
plot_daily_snotel_steam(snotel_df = wc_snotel, snotel_name = "Whiskey Creek", 
                        streamflow_df = all_streams_ne, stream_name = "Wheatfields")
plot_daily_snotel_steam(snotel_df = wc_snotel, snotel_name = "Whiskey Creek", 
                        streamflow_df = all_streams_ne, stream_name = "Tsaile Creek")
plot_daily_snotel_steam(snotel_df = wc_snotel, snotel_name = "Whiskey Creek", 
                        streamflow_df = all_streams_ne, stream_name = "Lukachai Creek")

```


### Compare SNOTEL peak SWE rank order with yearly summed streamflow rank order

```{r, echo = FALSE}

# whiskey Creek
wc_snotel_rank <- wc_snotel %>% 
  select("Date" = date, swe_m) %>% 
  addWaterYear() %>% 
  filter(waterYear > 2009) %>% 
  group_by(waterYear) %>% 
  summarize(max_swe_m = max(swe_m, na.rm = TRUE)) %>% 
  arrange(max_swe_m) %>% 
  mutate(rank_swe = 1:n())

ggplot(wc_snotel_rank, aes( x = as.character(waterYear), y = max_swe_m)) +
  geom_col() +
  labs(
    x = "Water Year",
    y = "Max SWE (m)",
    title = "Whiskey Creek Max SWE each Water Year"
  )

# beaver springs
bs_snotel_rank <- bs_snotel %>% 
  select("Date" = date, swe_m) %>% 
  addWaterYear() %>% 
  filter(waterYear > 2009) %>% 
  group_by(waterYear) %>% 
  summarize(max_swe_m = max(swe_m, na.rm = TRUE)) %>% 
  arrange(max_swe_m) %>% 
  mutate(rank_swe = 1:n())

ggplot(bs_snotel_rank, aes( x = as.character(waterYear), y = max_swe_m)) +
  geom_col() +
  labs(
    x = "Water Year",
    y = "Max SWE (m)",
    title = "Beaver Springs Max SWE each Water Year"
  )
```

```{r, echo = FALSE}
# peak discharge each water year
peak_discharge_wy <- all_streams_ne %>% 
  group_by(station, WY) %>% 
  summarize(discharge_cmps = max(discharge_cmps, na.rm = TRUE)) %>% 
  filter(discharge_cmps != "-Inf")

ggplot(peak_discharge_wy, aes(x = WY, y = discharge_cmps)) +
  geom_col(aes(fill = station)) +
  facet_wrap(~station) +
  labs(
    x = "Water Year",
    y = "Total Discharge (m3/s)",
    title = "Peak Discharge each Water Year"
  ) +
  theme_bw() 

```

### Compare peak discharge rank to snotel peak swe rank

```{r, include = FALSE}
# peak discharge ranking from WY 2010 for comparison to SNOTEL
peak_discharge_wy_10 <- peak_discharge_wy %>% 
  filter(WY > 2009) %>% 
  mutate(rank_stream = 1:n()) %>% 
  ungroup()

peak_discharge_wy_10_kt <- peak_discharge_wy_10 %>% 
  select(rank_stream, station) %>% 
  mutate(station = as.factor(station))

kruskal.test(rank_stream ~ station, data = peak_discharge_wy_10_kt)
# ranks are equal between each other

```


```{r, include = FALSE}
# run a mann_whitney test for each stream station to snotel station
mw_stream_snotel = function(snotel_df, stream_df, stream_name = ""){
  
  stream_indiv <- stream_df %>% 
    filter(station == stream_name)
  
  mann_u <- wilcox.test(stream_indiv$discharge_cmps, snotel_df$max_swe_m)
  
  return(mann_u)
}

### Whiskey Creek ###

# Asaayi and whiskey creek snotel
mw_stream_snotel(snotel_df = wc_snotel_rank, stream_df = peak_discharge_wy_10, stream_name = "Asaayi")

# Whiskey creek stream and whiskey creek snotel
mw_stream_snotel(snotel_df = wc_snotel_rank, stream_df = peak_discharge_wy_10, stream_name = "Whiskey Creek")

# Captain Tom Wash and whiskey creek snotel
mw_stream_snotel(snotel_df = wc_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Captain Tom Wash")

# Wheatfields and whiskey creek snotel
mw_stream_snotel(snotel_df = wc_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Wheatfields")

# Tsaile Creek and whiskey creek snotel
mw_stream_snotel(snotel_df = wc_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Tsaile Creek")

# Lukachai Creek and whiskey creek snotel
mw_stream_snotel(snotel_df = wc_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Lukachai Creek")

### Beaver Springs ####

# Asaayi and Beaver Springs snotel
mw_stream_snotel(snotel_df = bs_snotel_rank, stream_df = peak_discharge_wy_10, stream_name = "Asaayi")

# Whiskey creek stream and Beaver Springs snotel
mw_stream_snotel(snotel_df = bs_snotel_rank, stream_df = peak_discharge_wy_10, stream_name = "Whiskey Creek")

# Captain Tom Wash and Beaver Springs snotel
mw_stream_snotel(snotel_df = bs_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Captain Tom Wash")

# Wheatfields and Beaver Springssnotel
mw_stream_snotel(snotel_df = bs_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Wheatfields")

# Tsaile Creek and Beaver Springs snotel
mw_stream_snotel(snotel_df = bs_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Tsaile Creek")

# Lukachai Creek and Beaver Springs snotel
mw_stream_snotel(snotel_df = bs_snotel_rank, stream_df = peak_discharge_wy_10, 
                 stream_name = "Lukachai Creek")

```

The ranks of peak discharge per water year and maximum SWE are **not** statistically different for Whiskey Creek Snotel and Asaayi and Whiskey Creek Stream 

The ranks of peak discharge per water year and maximum SWE **are** statistically different for Whiskey Creek Snotel Captain Tom Wash, Wheatfields, Tsaile, Lukachai

The ranks of peak discharge per water year and maximum SWE are **not** statistically different for Beaver Springs Snotel and Asaayi and Whiskey Creek Stream 

The ranks of peak discharge per water year and maximum SWE **are** statistically different for Beaver Springs Snotel Captain Tom Wash, Wheatfields, Tsaile, Lukachai

